### 2020-06-12 UQonSets.tex
Updates from Overleaf

We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
of $\mes(\es)$ through its moments.

\medskip

Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
\begin{equation*}
\begin{split}
\emv
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ to
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
%
Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
solely an integral on $\domain$ and can be expanded in our present settings as follows
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%
Next we investigate uncertainty reduction criteria related to those indicators.