### 2020-06-12 appendix.tex

\begin{propo}
    \label{propo1}
%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for 
any $r\geq 1$,
\begin{equation*}
\begin{split}
\mathbb{E}[\mes(\es)^r]
&=\int_{D^{r}} \jointExcuProb
\productMeasure
,
\end{split}
\end{equation*}

where the product measure is denoted as
$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$.
%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
Here $\gp$ is defined on $D$, and for
$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
\medskip

In the particular case where $\gp$ is a multivariate Gaussian random field
%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively,
we have
\begin{align*}
\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
\end{align*}
where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ 
and covariance matrix $\covUU$, respectively defined blockwise by
%where both are defined blockwise by
\begin{align*}
\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
\in \mathbb{R}^{\no r}, \\
\text{and } \covUU &= \begin{pmatrix}
\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
\gp[u^{(r)}])\\
\vdots & & \vdots\\
\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
\gp[u^{(r)}])\\
\end{pmatrix}\in \mathbb{R}^{pr\times pr},
\end{align*}
each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times 
\no$.
%dimensional Gaussian random vector.
Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no 
r$-dimensional Gaussian probability density function
$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as
%$\varphi_{\no \times r}$ as
%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely
%cumulative distribution function, namely
\begin{equation*}
\begin{split}
&\jointExcuProb
=
\int_{T^r} \varphi_{\no  r}\left(\bm{v};
    ~\meanUU, \covUU\right)
    \mathrm{d}\bm{v},
\end{split}
\end{equation*}
In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
the latter probability directly writes in terms of the multivariate Gaussian
cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
this time by the way without requiring $\covUU$ %the latter
to be non-singular:
\begin{equation*}
\begin{split}
\jointExcuProb
&=
\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
\end{split}
\end{equation*}
where we have used the notations
$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and
$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no})
\in \R^{\no r}$.
\end{propo}
\begin{proof}
That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem
relying on the joint measurability of
$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$,
%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.
itself inherited from the assumed measurability for
$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that
\begin{equation*}
\begin{split}
\mathbb{E}[\mes(\es)^{r}]
&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
        \int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)})
\right) \right] \\
&
=
%\mathbb{E}\left[
%\int_{D^{r}}
%\prod_{i=1}^{r}
%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
%\productMeasure
%\right]
%=
\mathbb{E}\left[
\int_{D^{r}}
\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
\productMeasure
\right]
%\\
%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
%&
=\int_{D^{r}}
\jointExcuProb
\productMeasure.
\end{split}
\end{equation*}
The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of
$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$.
\end{proof}

The propositions below provide formulae for computations of expectations of moments of multivariate gaussian CDFs.

\begin{propo}
    \label{propo2}
Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
and $\covN, \covV$ be two covariance matrices in
$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
\begin{equation*}
\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
=
\varPhi_{ph}
\left(
    \bm{a}
;~
\bm{\Sigma}
\right),
\end{equation*}
where the vector $\bm{a} \in \R^{p h}$ is defined as
$\bm{a} := 1_h\otimes a = 
\left(a, \dots , a
\right)'$
 and the $p h\times p h$ covariance matrix is given by
 $\bm{\Sigma} := 
1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
\end{propo}

\begin{remark}
In blockwise representation, $\bm{\Sigma}$ can be expressed as follows:
\begin{align*}
% \bm{\Sigma}&=
\begin{pmatrix}
    \covN & &\\
        & \ddots &\\
        &   & \covN
\end{pmatrix}
+
\begin{pmatrix}
B\covV B' & \dots & B\covV  B'\\
\vdots & & \vdots\\
B\covV B' & \dots & B\covV B'\\
\end{pmatrix}
\end{align*}
\end{remark}

\begin{proof}
By definition of $\Phi_{p}$, for $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
$$
\mathbb{P}(N\leq a + BV | V)
=
\varPhi_{p}\left( a + BV; \covN \right).
$$
Now for $\varPhi_{p}\left( a + BV; \covN \right)^h$, provided that the probability space is sufficiently large to accomodate $h$ independent Gaussian random vectors $N_i\sim \mathcal{N}_{p}(0,\covN)$ (which is silently assumed here), using the former equality delivers
$$
\varPhi_{p}\left( a + BV; \covN \right)^h
=
\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V).
$$
Now by independence of the $N_i$'s we obtain the joint conditional probability
$$
\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V)
=
\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V),
$$
whereof, by virtue of the law of total expectation,
\begin{equation*}
\begin{split}
\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
&=\mathbb{E}\left[\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V)\right]\\
&=\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV)\\
&=\mathbb{P}(W_1 \leq a, \dots, W_h\leq a)\\
&=\varPhi_{ph}
\left(
1_{h} \otimes a
;
(1_{h}1_{h}')\otimes (B\Sigma_{V} B') + 
I_{h}\otimes \covN
\right),
\end{split}
\end{equation*}
where $\mathbf{W}=(W_1,\dots,W_h)$ with $W_i=N_i- BV \ (1\leq i \leq h)$
and the last line follows $\mathbf{W}$ forming a Gaussian vector (by global independence of the $N_i$'s and $V$) and from the definition of $\varPhi_{p h}$. The covariance matrix $\mathbf{\Sigma}$ of $\mathbf{W}$ is obtained by noting that $\operatorname{cov}(W_i,W_j)=B \covV B' + \delta_{ij} \covN \ 
(i,j \in \{1,\dots,h\})$.
%the latter concatenated Gaussian vector is obtained by straightfoward calculation. 
\end{proof}

We now generalize Proposition~\ref{propo2} to the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector.
%Namely, for monomials of degree $k$, we have.

\begin{propo}
    \label{propo3}
Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$. Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
    \begin{equation}
    \mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
    =
\varPhi_{p H}
\left(
    \bm{a}
;
\mathbf{\Sigma}
\right),
\end{equation}
with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
\begin{equation}
\Sigma_{i,j}=
(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
\end{equation}
\end{propo}

\begin{remark}
Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
\begin{equation*}
%    \bm{\Sigma} = \begin{pmatrix}
%        \Sigma_{11} & & \Sigma_{1g}\\
%        & \ddots &\\
%        \Sigma_{g1}&   & \Sigma_{gg}
%\end{pmatrix},~
\Sigma_{ij} =
\begin{pmatrix}
B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
\vdots & & \vdots\\
B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
\end{pmatrix}
+
\delta_{ij}
\begin{pmatrix}
    \covN_i & &\\
        & \ddots &\\
        &   & \covN_i
\end{pmatrix}%\in\mathbb{R}^{lh_i\times lh_j}
\end{equation*}
Here each $\Sigma_{ij}$ is made of $h_i$ times $h_j$
(vertically/horizontally) $p \times p$ sub-blocks, hence possesses $ph_i$ lines and $ph_j$ columns.
\end{remark}

\begin{proof}
    The proof relies (again) heavily on the fact that, by definition of $\Phi_{p}$, for any covariance matrix $\covN \in \R^{p \times p}$, $a\in \R^p$, $B\in \R^{p \times q}$, and $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
    $$
    \mathbb{P}(N\leq a + BV | V)
    =
    \varPhi_{p}\left( a + BV; \covN \right).
    $$
In particular, for globally independent $N_{i,j} \sim \mathcal{N}_{p}(0_{p},\covN_i)$ $(1\leq j \leq h_i, 1\leq i \leq g)$,
\begin{equation*}
\begin{split}
\prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i}
&=
\prod_{i=1}^{g}
\prod_{j=1}^{h_{i}}
\mathbb{P}(N_{i,j}\leq a_i + B_i V | V)\\
&=\mathbb{P}(N_{1,1}\leq a_1 + B_1 V, \dots, N_{g,h_{g}}\leq a_g + B_g V | V),
\end{split} 
\end{equation*}
so that, by the law of total expectation,
    \begin{equation*}
    \begin{split}
    \mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
=
\mathbb{P}(W_{1} \leq 1_{h_{1}} \otimes a_1, \dots, W_{g} \leq 1_{h_{g}} \otimes a_g)
    \end{split}
    \end{equation*}
where
$W_{1}=(N_{1,1}- B_1 V, \dots, N_{1,h_{1}}- B_1 V), 
W_{2}=(N_{2,1}- B_2 V, \dots, N_{2,h_{2}}- B_2 V), 
\dots, W_{g}=(N_{g,1}- B_g V, \dots, N_{g,h_{g}}- B_g V)$. Noting that $\mathbf{W}=(W_1,\dots, W_g)$ is a centred $p H$-dimensional Gaussian random vector, we finally obtain that
    \begin{equation*}
\begin{split}
\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
=
\varPhi_{p H}\left(\bm{a};\mathbf{\Sigma}\right),
    \end{split}
\end{equation*}
with $\bm{a}=(1_{h_{1}} \otimes a_1, \dots, 1_{h_{g}} \otimes a_g)$ and $\bm{\Sigma}=(\operatorname{cov}(W_i,W_j))_{i,j \in \{1,\dots, g\}}$.
\end{proof}

Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\ibv$ (Proposition \ref{propo_eibv}) and on the $\emv$ (Proposition \ref{propo_emv}) for which we provide proofs below.

\begin{proof}{(Proposition \ref{propo_eibv})}
Applying Tonelli-Fubini followed by the law of total
expectation first delivers
\begin{equation*}
\begin{split}
\eibv_{[n]}(\bm{x})
&=\int_{D}
\currentExp{\futureProba{\gp[\uu]\in
        T}(1-\futureProba{\gp[\uu]\in T})} d\mes(u) \\
&=\int_{D} \varPhi_{\no}\left(\bt;
~\futureMean{\uu},
\futureCov{u,u}\right) d\mes(u)\\
&-\int_{D} \currentExp{
    \varPhi_{\no}\left(\bt;
    ~\futureMean{\uu},
    \futureCov{u,u}\right)^2
}
d\mes(u), 
\end{split}
\end{equation*}
%
where $\futureCov{u,u}$ denotes the $\no \times \no$ covariance matrix between all $\no$
responses at point $u$ conditional on the first $n+1$ observation batches.
Now, by using co-kriging update formulae and our shortcut notation for the CDF of centred
multivariate Gaussian vectors, we observe that
\begin{equation*}
\begin{split}
%\mathbb{E}_{n}[\left(
&\varPhi_{\no}\left(\bt;~\futureMean{\uu}, \futureCov{u, u}\right) 
%\right)^2]
\\
=&
\varPhi_{\no}\left(\bt-\futureMean{\uu}; \futureCov{u, u}\right) \\
=&
\varPhi_{\no}\left(\bt-\currentMean{\uu}-\lambda_{[n+1,n+1]}(u)^T(\gp[\bm{x}_{n+1}]
-\currentMean{\bm{x}_{n+1}}), \futureCov{u, u}\right) \\
=&
\varPhi_{\no}\left(a + BV, \futureCov{u, u}\right),
\end{split}
\end{equation*}
with $a=\bt-\currentMean{\uu}$, %+\lambda_{[n+1,n+1]}(u)^T\currentMean{\bm{x}_{n+1}}$,
$B=-\lambda_{[n+1,n+1]}(u)^T$
and $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}}$.
Applying Proposition~\ref{propo2} then delivers that
%
\begin{equation*}
\begin{split}
&\currentExp{
    \varPhi_{\no}\left(\bt;~\futureMean{\uu}, \futureCov{u, u}\right)^2 
}
%\\
=\varPhi_{2\no}
\left(
\left(
\begin{matrix}
\bt-\currentMean{\uu}\\
%\vdots \\ 
%\vdots \\ 
%\vdots \\ 
\bt-\currentMean{\uu}
\end{matrix}
\right);
\mathbf{\Sigma}_{[n]}(\uu)
\right),
\end{split}
\end{equation*}
with $\mathbf{\Sigma}_{[n]}(\uu)$ as in the formulation of the proposition. This completes the proof.
\end{proof}

\begin{proof}{(Proposition \ref{propo_emv})}
\begin{equation*}
\begin{split}
\eemv_{[n]}(\bm{x})
%\operatorname{Var}[\mes(\es)]
&=\int_{D^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\currentExp{\left( \int_{D} \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right) d\mes(u) \right)^2}.
\end{split}
\end{equation*}
The second term can be worked out in semi-analytical form by writing it as
%
\begin{equation*}
\begin{split}
%\eemv_{[n]}(\bm{x})
%\operatorname{Var}[\mes(\es)]
&\int_{D^2} 
\currentExp{
\varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
\varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
}
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
=&
\int_{D^2} 
\currentExp{
    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
}
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v),\\
\end{split}
\end{equation*}
and expanding the integrand (for fixed $n$, $\uu, \vv$) as
\begin{equation*}
\begin{split}
&
\currentExp{
    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
}\\
=&
\currentExp{
    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
}
\end{split}
\end{equation*}
with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
\end{proof}