### 2020-06-23 main_new.tex
updates

\documentclass[aoas]{imsart}

\input{commands.tex}
\usepackage{amsmath}
\usepackage{pstricks,pst-grad}
\usepackage{graphicx}
\usepackage{floatrow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\floatsetup[table]{capposition=top}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}                     % horizontal lines in tables
\usepackage{comment}

% == Enable text degreehttps://www.overleaf.com/project/5a3268379ecbdc657d8767e8
\usepackage{textcomp}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% == Trygve Test
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

% put your definitions there:
\startlocaldefs
\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
\endlocaldefs

% New Marcos.
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cleveref}
\input{macros}

\begin{document}

\begin{frontmatter}

% "New Title of the paper"
\title{Autonomous Oceanographic Data Collection Estimating Excursion Sets of Vector-valued Gaussian Random Fields} 
\runtitle{Autonomous Oceanographic Data Collection}

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}

\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e2}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e4}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e5}}

\runauthor{TO. Fossum et al.}
\end{aug}

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for marine science and maritime resource management. Faced with
  limited resources in understanding processes in the water-column,
  the combination of statistics and autonomous systems provide new
  opportunities for experimental design. In this work we develop
  methods for efficient spatial sampling applied to the mapping of
  coastal ocean processes by providing informative descriptions of
  spatial characteristics of ocean phenomena. Specifically, we define
  a design criterion based on uncertainty in the excursions of
  vector-valued Gaussian random fields, and derive tractable
  expressions for the expected Bernoulli variance reduction in such a
  framework. We demonstrate how this criterion can be used to
  prioritize sampling efforts at locations that are ambiguous, making
  exploration more effective. We use simulations to study the
  properties of methods and to compare them with state-of-the-art
  approaches, followed by results from field deployments with an
  autonomous underwater vehicle (AUV) as part of a study mapping the
  boundary of a river plume. The results demonstrate the potential of
  combining statistical methods and robotic platforms to effectively
  inform and execute data-driven environmental sampling.

%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions. These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.

\end{abstract}

\begin{keyword}
\kwd{Ocean Sampling}
\kwd{Excursion Sets}
\kwd{Gaussian Processes}
\kwd{Experimental Design}
\kwd{Autonomous robots}
\kwd{Adaptive Information Gathering}
\end{keyword}

\end{frontmatter}
\section{Introduction}

%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.

Motivated by the challenges related to efficient data collection
strategies for our vast oceans, we combine spatial statistics, design
of experiments and marine robotics in this work.  The
multidisciplinary efforts enable information-driven data collection in
regions of high-interest.

\subsection{Oceanic data collection and spatial design of experiments}

Monitoring the world's oceans has gained increased importance in light
of the changing climate and increasing anthropogenic impact. Central
to understanding the changes taking place in the upper water-column is
knowledge of the bio-geophysical interaction driven by an
agglomeration of physical forcings (e.g. wind, topography, bathymetry,
tidal influences, etc.) and incipient micro-biology driven by
planktonic and coastal anthropogenic input, such as pollution and
agricultural runoff transported into the ocean by rivers.  These often
result in a range of ecosystem-related phenomena such as blooms and
plumes, with direct and indirect effects on society. \kc{we should
  look for a citation here. Check with John Ryan} One of the
bottlenecks in the study of such phenomena lies however in the lack of
observational data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations in which ocean processes transpire, prompting the need for
effective means of data collection.  By \emph{sampling}, we refer here
primarily to the design of observational strategies in the spatial
domain with the aim to pursue measurements with high scientific
relevance.  Spatial statistics and methods for experimental design can
clearly contribute to this sampling challenge.

%applications and beyond.  
 %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance. This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.

Data collection at sea has typically been based on static buoys,
floats, or ship-based methods, with significant logistical limitations
that directly impact coverage and sampling resolution. Modern methods
using satellite remote-sensing provide large-scale coverage but have
limited resolution, are limited to sensing the surface, and are
impacted by cloud cover. Numerical ocean models similarly find it
challenging to provide detail at fine scale \citep{Lermusiaux:2006},
and also come with computational costs that can be limiting. The
advent of robust mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling in the ocean (Fig. \ref{fig:envir1}). In particular,
autonomous underwater vehicles (AUVs) have advanced the state of data
collection and consequently have made robotics an integral part of
ocean observation \citep{das11b,Das2015,fossuminformation,fossum18b}.

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of a range of ocean sensing opportunities.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on 
    ship-based sampling has been augmented by autonomous
    robotic vehicles. % and their interactions.  
    AUV platforms are an integral part of this network being able to
    reason and make decisions for efficient onboard adaptive sampling.
    % using the sense-plan-act control approach to autonomous control. 
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective oceanographic observation.}
  \label{fig:envir} \end{figure}

Surveys with AUVs are usually limited to observations along fixed
transects that are pre-scripted in mission plans created manually by a
human operator. Missions can be specified operating on a scale of
hundreds of meters to tens of kilometers depending on the scientific
context. Faced with limited coverage capacity, a more effective
approach is to instead use onboard algorithms to continuously
evaluate, update, and refine future sampling locations, making the
data collection \emph{adaptive}.  In doing so, the space of sampling
opportunities is still limited by a waypoint graph, which forms a
discretization of the search domain where the AUV can navigate;
however the AUV can now modify its path at each waypoint based on
in-situ measurements and calculations onboard using onboard
deliberation \citep{py10,Rajan12,Rajan12b}.  Full numerical ocean
models based on complex differential equations cannot be run onboard
the AUV with limited computational capacity, and statistical models
relying on random field assumptions are relevant as a means to
effectively update the onboard model from in-situ data, and to guide
AUV data collection trajectories.

The work presented here is primarily inspired by a case study
pertaining to using an AUV for spatial characterization of a frontal
system generated by a river plume. Fig. \ref{fig:nidelven} shows the
survey area in Trondheim, Norway, where cold freshwater enters from
the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coriolis force the
cold fresh water tends to flow east. Depending on the variations in
river discharge, tidal effects, coastal current and wind, this
boundary often gets distorted, and knowledge about its location is
highly uncertain, making deterministic planning challenging. The goal
is therefore, to use AUV measurements for improved description of the
interface between fresh and oceanic waters.  It is often not possible
to sample the biological variables of fundamental interest in such AUV
operations, but off-the-shelf instruments provide temperature and
salinity measurements which serve as proxys for the underlying
biological phenomenon. With the help of a vector-valued random field
model for temperature and salinity, one can then aim to describe the
plume.  The goal of plume characterization, in this way, relates to
that of estimating some regions of the domain, typically excursion
sets (ESs), when implicitly defined by the vector-valued random field.
In our context of environmental sampling, the joint salinity and
temperature excursions of a river plume help characterize the
underlying bio-geochemical processes
\citep{hopkins2013detection,Pinto2018}. Motivating examples for ESs of
multivariate processes are also abundant in other contexts, for
instance in medicine, where doctors do not rely solely on a single
symptom but must see several combined effects before making a
diagnosis.

The questions tackled here hence pertain to the broader area of
spatial data collection and experimental designs for vector-valued
random fields.
%Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs. 
Given the operational constraints on AUV movements and the fact that
surveys rely on successive measurements along a trajectory, addressing
corresponding design problems calls for sequential strategies.  Our
main research angle in the present work is to extend sequential design
strategies from the world of spatial statistics and computer
experiments to the setting of both vector-valued observational data
and experimental designs for feasible robotic trajectories. We
leverage and extend recent progress in expected uncertainty reduction
for ESs of Gaussian random fields (GRFs) in order to address this
research
problem. %approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
We briefly review recent advances in targeted sequential design of
experiments based on GRFs before detailing other literature related to
AUV sampling and our contributions prior to outlining the rest of the
paper.

\subsection{Random field modeling and targeted sequential design of experiments}

While random field modeling has been one of the main topics throughout
the history of spatial statistics \citep{Krige1951a,Stein1999}, even
for vector-valued random field models with associated prediction
approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003},
there has lately been a renewed interest for random field models in
the context of static or sequential experimental design, be it in the
context of spatial data collection \citep{Mueller2007} or in
simulation experiments \citep{Santner.etal2003}. As detailed in
\cite{Ginsbourger2018}, GRF models have been used in particular as a
basis to sequential design of simulations dedicated to various goals
such as global optimization and set estimation
\citep{Adler.Taylor2007}. Of particular relevance to our context,
\cite{Bect.etal2012} focuses on strategies to reduce uncertainty in
volumes of excursion exceeding a prescribed threshold, while
\cite{chevalier2014fast} contributes a computationally efficient
implementation of sequential strategies. Rather than focusing on
excursion volumes, approaches were investigated in
\cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
with ambitions of estimating sets themselves. Recently, sequential
designs of experiments for the conservative estimation of ESs based on
GRF models were presented in \citep{Azzimonti.etal}.

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be used efficiently for
assimilating sequential data \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, the PhD thesis \citep[][p.82]{stroh} mentions general
possibilities of stepwise uncertainty reduction strategies for ES's in
the context of designing simulations, but the outputs are mainly
assumed independent.

% From earlier version; to be relocated?
%Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes. 
\subsection{Previous work in AUV sampling}

Other statistical work in the oceanographic domain include
\cite{wikle2013modern} focusing on hierarchical statistical models,
\cite{sahu2008space} studying spatio-temporal models for sea surface
temperature and salinity data and \cite{mellucci2018oceanic} looking
at the statistical prediction of features using an underwater glider.
In this paper the focus is not on statistical modeling, but rather on
statistical properties of computation for efficient data
collection. In this setting, we combine novel possibilities in marine
robotics with spatial statistics and experimental design to provide
useful AUV sampling designs.

Adaptive in-situ AUV sampling of an evolving frontal feature has been
explored in \cite{fronts11,Smith2016,Pinto2018,costa19}. These
approaches typically use a reactive-adaptive scheme, whereby
exploration does not rely on a statistical model of the environment,
but rather adaptation is based on closing the sensing and actuation
loop. Myopic sampling, i.e. stage-wise selection of the path (on the
waypoint graph), has been used for surveys
\citep{singh2009efficient,Binney2013} that focus largely on reducing
predictive variance or entropy. These criteria are widely adopted in
the statistics literature on spatio-temporal design as well
\citep{bueso1998state,zidek2019monitoring}. However, variance and
entropy reduction are independent of the actual data realizations
under the assumptions of GP models, so it has limited flexibility for
active adaptation of trajectories based on what is measured.  The use
of data-driven adaptive criteria was introduced to include more
targeted sampling of regions of scientific interest in \cite{Low2009}
and \cite{fossuminformation}.

The primary contributions of this work are:

\begin{itemize}
\item Extending expected uncertainty reduction criteria to
  vector-valued settings.  
\item Closed form expressions for the expected integrated Bernoulli
  variance (IBV) of the excursions in GRFs. 
\item Algorithms for myopic and multiple-step ahead sequential
  strategies for optimizing AUV sampling with respect to the mentioned
  criteria. 
\item Replicable experiments on synthetic test cases with accompanying
  code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
\item Results of field trials running myopic strategies onboard an AUV
  for the characterization of a river plume. 
\end{itemize}

The remainder of this paper is organized as follows:
%Section \ref{sec:bg} provides select background on ocean sampling. 
Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and
the design criteria connected to the IBV for excursions of
vector-valued GPs. Section \ref{sec:heuristics} builds on these
assumptions when deriving the sequential design criteria for adaptive
sampling. In both sections properties of the methods are studied using
simulations. Section \ref{sec:case_study} demonstrates the methodology
used in field work characterizing a river plume. Section
\ref{sec:concl_disc} contains a summary and a discussion of future
work.
%\input{section2}

\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
\label{sec:ESEP}

Section \ref{sec:bg_and_notation} introduces notation and co-Kriging
equations of multivariate GRFs.  Section \ref{sec:set_uq} presents
uncertainty quantification (UQ) techniques on ESs of GRFs, in
particular the IBV and the excursion measure variance (EMV).  Section
\ref{sec:eibv} turns to the effect of new observations on EMV and IBV,
and semi-analytical expected EMV and IBV over these observations are
derived.
%The resulting formulae form the backbone of the expected sequenatil uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate
example relevant for sampling temperature and salinity in our domain.

\subsection{Background, Notation and Co-Kriging}
\label{sec:bg_and_notation}

We denote by $\gp$ a vector-valued random field indexed by some
arbitrary domain $\domain$, and assume values of the field at any
fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a
$\no$-variate random vector ($\no\geq 2$). In the river plume
characterization case, $\domain$ is a prescribed domain in
Trondheimsfjord, Norway (for the purpose of our AUV application, a
discretization of a $2$-dimensional domain at fixed depth is
considered), and $\no=2$ with responses of temperature and salinity. A
bivariate GRF model is assumed for $\gp$. To motivate concepts,
Fig.~\ref{example_excu} (left and middle \kc{If you split this figure
  in its parts and add labels, it would be more elegant and precise to
  explicitly refer to them.}) shows a realization of such a
vector-valued GRF on $\domain=(0,1)^2$. Fig.~\ref{example_excu} (right
\kc{See above}) represents a by-product of interest derived from these
realizations, namely regions i) in red, where both temperature and
salinity are high (i.e., exceeding respective thresholds), indicative
of ocean water ii) in white, where both temperature and salinity are
low, indicative of riverine water, and iii) in pink, where one
variable is above and the other below their respective thresholds,
indicative of mixed waters.

\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/example_excu_4.png}
  \caption{Realization of a bivariate GRF and excursion set above some
    threshold ({\kc{JE: are means around 5 and 30 here, so same as
        later in the example?}} ). Joint excursion in red, excursion
    of a single variable in pink.}
\label{example_excu}
\end{figure}

%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
For the general setting of a $\no$-variate random field, we are
interested in recovering the set of locations $\es$ in the domain for
which the components of $\gp$ lie in some set of specified values
$\T\subset \mathbb{R}^{\no}$; in other words \textit{the pre-image of
  $T$ by $\gp$}:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
%
%Our goal here is to develop approaches to quantify and improve the
%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}

If we assume that $\gp$ has continuous trajectories 
%(almost surely \kc{This seems very handwavy; would delete}) 
and $T$ is closed, then $\es$ becomes a Random Closed Set
\citep{Molchanov2005} and concepts from the theory of random sets will
prove useful to study $\es$.
%under some measure defined on the domain.
%\textcolor{red}{Is this really needed?} 
Note that while some aspects of the developed approaches do not call
for a specific form of $\T$, we will often, for purposes of
simplicity, stay with the case of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

\medskip

%\subsubsection{Notation}
%In order to work at this level of generality, some notational tricks are needed. Indeed, when considering a new observation, one shall in general specify which component of the field was observed and where it was observed. This leads us to introduce the concept of \textit{generalized location}.

Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of
$\gp[\spatloc]$ ($1\leq \ell\leq \no$), we use the term
\textit{generalized location} for the couple $x=(\spatloc,\ell)$.
%Given such a generalized location notation $x$,  
The notation $\gp[x]$ will be used to denote $\gp[\spatloc,\ell]$ 
and
%
%This slight change of notation 
will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging. %Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$. In the following, the letter $x$ will be reserved for generalized locations, while 
The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
%
% Replacement by overlooked revised chunks from June 19th
%\begin{comment}
%Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
%\begin{align*}
%\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
%\end{align*}
%In general,  corresponding to batches of observations. 
%The field values of batch observations are denoted by
%\begin{align*}
%\gp[\bm{x}]:=
%\left(\gp[\spatloc_1,\ell_1], ...,
%\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
%\end{align*}
%In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.

%Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance . This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
%We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).

%Assuming that one wishes 
%to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
%given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
%The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
%
%\begin{equation}\label{eq:cokrig_mean}
%\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
%\end{equation}
%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
%where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations. 
%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
%The associated co-kriging %(conditional) 
%residual cross-covariance function can be expressed in the same vein via
%
%\begin{equation}\label{eq:cokrig_cov}
%k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
%\end{equation}

%\subsubsection{Co-kriging update formulae}

%In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations. Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case. Because of our representation of co-kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}. The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.

%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$. The mean update formulae is
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
% delivers that
% 
%\begin{equation}\label{eq:meanCoK}
%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
%\end{equation}
%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
%The updated co-Kriging residual cross-covariance function is
%\begin{eqnarray}\label{eq:varCoK}
%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
%\end{eqnarray}
%\medskip
%
%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
%\end{comment}
%
%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
%one for a single observation, provided some notation adjustments are made. This motivates the following.
Furthemore, boldface letters will be used to denote concatenated
quantities corresponding to batches of observations.  Given a dataset
consisting of
$q$ observations at spatial locations
$\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in
\domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in
\lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
we use the concatenated notation
\begin{align*}
\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
\end{align*}
We also compactly denote the field values at those different locations by
\begin{align*}
\gp[\bm{x}]:=
\left(\gp[\spatloc_1,\ell_1], ...,
\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
\end{align*}
%
For a second order random field $(Z_{\spatloc})_{\spatloc \in
  \domain}$ with mean $\mu$ and matrix covariance function $K$,
$\mu$ is naturally extended to $\domain \in \lbrace 1, ...,
\no\rbrace$ into a function of $x=(\spatloc,
\ell)$ and is further straightforwardly vectorized into a function of
$\bm{x}$. As for $K$, it induces a covariance kernel
$k$ on the set of extended locations via $k((\spatloc,
\ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell,
  \ell'}$. In vectorized/batch form, $k(\bm{x},
\bm{x}')$ then amounts to a matrix with numbers of lines and columns
the numbers of generalized locations in the batches
$\bm{x}$ in
$\bm{x}'$, respectively \kc{This last sentence doesn't parse.}. Such
vectorized quantities turn out to be useful in order to arrive at
simple expressions for co-Kriging equations below.
%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}

%\medskip

Given a GRF $\gp$ and observations of some of its components at some
points in the domain, one can predict the value of the field at some
unobserved location $\spatloc\in \domain$ by using the conditional
mean of $\gp[\spatloc]$, conditional on the data \kc{...conditional
  mean on the data?}. This coincides with (co-)kriging equations,
which tell us precisely how to compute conditional means and
covariances conditional on an arbitrary dataset.  We will present a
general form of co-kriging, in the sense that it allows inclusion of
several observations (batch) at a time; observations at a given
location $u \in \domain$ may only include a subset of the components
of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).

Assuming that $n$ batches of observations are available with sizes
$q_1,\dots, q_n$, and that one wishes to predict $\gp[\bm{x}]$ for
some batch of $q\geq 1$ generalized locations
$\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
the simple cokriging mean then amounts to kriging with respect to a
scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
%
\begin{equation}\label{eq:cokrig_mean}
\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
\end{equation}
where $\mu$ is $Z$'s initial mean function,
$\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
q_i$)-dimensional vector of observed (noisy) responses of
$Z$ at all considered generalized locations, and
$\lambda_{[n]}(\bm{x})$ is a vector of weights equal to
$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
$$
with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
\bm{x}_n)$ and where
$\Delta_{[n]}$ stands for the covariance matrix of
Gaussian-distributed noise assumed to have affected measurements up to
batch
$n$. The matrix in parenthesis will be assumed to be non-singular
throughout the presentation. The co-kriging %(conditional)
residual (cross-)covariance function (with respect to batches of
generalized locations) can also be expressed in the same vein via
%
\begin{equation}\label{eq:cokrig_cov}
k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T 
\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)
\lambda_{{[n]}}(\bm{x}').
\end{equation}

%\subsubsection{Co-kriging update formulae}

Let us now consider the case where a co-kriging prediction of $Z$ was
made with respect to $n$ batches $\bm{x}_i$ of generalized locations,
concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
and one wishes to update the prediction by incorporating a new vector
of observations $\mathbf{z}_{n+1}$ measured at a batch of
$q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
Thanks to our representation of co-kriging in terms of simple kriging
with respect to generalized locations, a strightforward adaptation of
the batch-sequential kriging update formulae from
\citep{Chevalier.etal2013a} suggests that
% 
\begin{equation}\label{eq:meanCoK}
\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
\end{equation}
where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
to the kriging weigths associated with the last $q_{n+1}$ responses
when predicting at $\bm{x}$ relying on all measurements until batch
$(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
Similarly, the updated co-kriging residual (cross-)covariance function
then becomes
\begin{align}\label{eq:varCoK}
k_{[n+1]}(\bm{x},\bm{x}') & = k_{[n]}(\bm{x},\bm{x}')\\
 \nonumber - & \lambda_{[n+1,n+1]}(\bm{x})^T 
\left(k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}\right)
%k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
\lambda_{{[n+1,n+1]}}(\bm{x}'),
\end{align}
%\medskip
%
As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
these update formulae naturally extend to Universal Kriging in
second-order settings and apply without Gaussian assumptions. We will
now see how the latter formulae are instrumental in deriving
semi-analytical formulae for stepwise uncertainty reduction criteria
for vector-valued random fields.

\subsection{Uncertainty Quantification on ESs of multivariate GRFs}
\label{sec:set_uq}

We now introduce quantities that allow UQ on the volume of the ES
$\es$. Let $\mes$ be a (locally finite, Borel) measure on
$\domain$. We want to investigate the probability distribution of
$\mes(\es)$ through its moments.  Centred moments may be computed
using Proposition~\ref{propo1} developed in the appendix.  In
particular, the excursion measure variance
$\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the excursion
probabilities (EP)

\begin{equation*}
\begin{split}
\emv
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ is
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}

where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative
distribution function (CDF), which can be calculated, using for
instance, the code of \cite{genz2009computation}.

Note that this quantity requires the solution of an integral over
$\domain^2$. In contrast, and as in the scalar-valued case, the IBV of
\cite{bect2019} involves solely an integral on $\domain$ and can be
expanded as
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%

\subsection{Expected IBV and EMV}
\label{sec:eibv}

We compute the expected effect of the inclusion of new observations on
the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the
same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
expectation and probability conditional on the first $n$ batches of
observations, respectively. We want to compute the effect of including
a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and
$\ibv$. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the
conditional law $\mathbb{P}_{\stage}$.

\medskip

In order to study the effect of the inclusion of a new data point, we
let $ \currentIBV(\bm{x}; \bm{y}) $ denote the expected IBV under the
current law of the field \kc{Is this a legitimate expression 'law of
  the field'?}, conditioned on observing $\bm{y}$ at $\bm{x}$
(generalized, possibly batch observation). The expected effect of a
new observation on the IBV is then shown by
\begin{equation}\label{def:eibv}
    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right]
\end{equation}
where $\bm{Y}$ is distributed according to the current law of
$Z_{\bm{x}}$ and with independent noise having covariance matrix
$\Delta_n$.

We next present a result that allows efficient computation of $\EIBV$
as an integral of CDFs of multivariate gaussians. This will prove
useful when designing sequential expected uncertainty reduction
strategies.

\begin{propo}
\label{propo_eibv}
\begin{equation}
\begin{split}
\currentEIBV(\bm{x})
&=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
&-\int_{D} \varPhi_{2\no}
\left(
\left(
\begin{matrix}
\bt-\currentMean{u}\\
\bt-\currentMean{u}
\end{matrix}
\right);
\mathbf{\Sigma}_{[n]}(\uu)
\right)
d\mes(u),
\end{split}
\end{equation}
where the matrix $\mathbf{\Sigma}_{[n]}(\uu)$ is defined as
\begin{equation*}
\begin{split}
\mathbf{\Sigma}_{[n]}(\uu)&=
\left(
\begin{matrix}
\currentCov{u, u} & \currentCov{u, u}-\futureCov{u, u}\\
\currentCov{u, u}-\futureCov{u, u} & \currentCov{u, u}
\end{matrix}
%\begin{matrix}
%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
%\Delta_{[n]}(\uu) & \currentCov{u, u}
%\end{matrix}
\right).\\
\end{split}
\end{equation*}
\end{propo}

As for the expected EMV, a similar result may be derived.
\begin{propo}
\label{propo_emv}

\begin{equation*}
\begin{split}
\eemv_{[n]}(\bm{x})
&=\int_{\domain^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\int_{\domain^2}\currentExp{
    \varPhi_{\no}\left(a_1+B_1 V; K_{[n+1]}(u)\right)
    \varPhi_{\no}\left(a_2+B_2 V; K_{[n+1]}(v)\right)
}\mathrm{d}\mes^{\otimes}(u, v),
\end{split}
\end{equation*}

with $V=\gp[\bm{x}]-\currentMean{\bm{x}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x},\bm{x}))$ and $a_1=\bt-\currentMean{\uu}$,
$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.

This integrand finally boils down to an evaluation of the
$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3}. Details
are omitted here for brevity as we do not implement this criterion
further and which requires numerical integration over $\domain^2$.
\end{propo}

We remark that Propositions \ref{propo_eibv} and \ref{propo_emv} are
twofold generalizations of results from \cite{chevalier2014fast}: they
extend previous results to the multivariate setting and also allow for
the inclusion of batch or heterotopic observations through the concept
of generalized locations.  A key element for understanding of these
propositions is that the conditional co-Kriging means entering in the
EPs depend linearly on (batch) observations. The conditional equality
expressions thus become linear combinations of Gaussian variables
whose mean and covariance are easily calculated.  Related closed-form
solutions have been noted in similar contexts
\citep{bhattacharjya2013value,stroh}, but not generalized to our
situation with random sets for vector-valued GRFs.

\subsection{Expected Bernoulli Variance on a two dimensional Example}
\label{Sec:UnivarEx}

We illustrate the EBV associated with different designs on a simple
bivariate example. This mimics our river plume application and hence
the first and second component of the random field will be called
\textit{temperature} and \textit{salinity} for illustrative
purpose. Effects of the hyperparameters of the GRF prior on the
excursion probabilities will also be studied. For simplicity, we begin
with a \textit{pointwise} example, considering a single bivariate
gaussian distribution (i.e. no spatiality).

\subsubsection{A pointwise study}

Say we want to study the excursion probability of a bivariate gaussian
above some threshold, where the thresholds are set equal to the mean;
$\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
salinity, and we play with the temperature and salinity correlation
and variances to study the effect on the EP and expected Bernoulli
variance.

\begin{figure}[!b] 
\centering
\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
\caption{Density contour plots with different correlations between
  temperature and salinity. The densities have unit variance and the
  thresholds are identical to the mean values $5^o C$ and
  $30$ g/kg.}
\label{illus_bivarDens}
\end{figure}

Fig. \ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity. The displayed densities have unit standard deviations for
both temperature and salinity, but we also study the effect of
doubling the standard deviations.

Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
Bernoulli variance (second row) for the examples indicated in Fig.
\ref{illus_bivarDens}. The EPs increase with the correlation as there
is a strong tendency to have concurrently low temperature and
salinity. The Bernoulli variance is similarly large for high
correlations. EPs and Bernoulli variances are the same for standard
deviation $1$ or $2$, which implies that high variability in
temperature and salinity is not captured in the $p(1-p)$ expression.

\begin{table}[!t] \centering \caption{EP and Bernoulli variance for
    different correlations and variances (top rows), and expected
    Bernoulli variances for both temperature and salinity data $\by$
    and temperature $y_2$ (bottom rows). \kc{$\mu_1$  or $\by$? And
      whatever happened to make the terms more accessible with $t$ and
    $s$ as we had before?}}
  \begin{tabular}{c|ccc|ccc}
 &\multicolumn{3}{c}{$\sigma_1=\sigma_2=1$} & \multicolumn{3}{c}{$\sigma_1=\sigma_2=2$} \\
\hline
Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
\hline
$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
EIBV, Temperature and Salinity & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
EIBV, Temperature only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
\hline
\end{tabular}
\label{tab:sim_rhoab}
\end{table}

The bottom two rows of Table \ref{tab:sim_rhoab} show results of
expected Bernoulli variance. This is presented for a design gathering
both data types, and for a design with temperature measurements
alone. When both data are gathered, the measurement model is
$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with
$\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$,
$\epsilon \sim N(0,0.5^2)$ when only temperature is measured.  For
this illustration, Table \ref{tab:sim_rhoab} shows that the expected
Bernoulli variance gets lower with larger standard deviations
$\sigma_1$ and $\sigma_2$. The reduction of Bernoulli variance is
largest for the cases with high correlation $\gamma$. Albeit smaller,
there is also uncertainty reduction when only temperature is measured
(bottom row), especially when temperature and salinity are highly
correlated. When correlation is low ($\gamma=0.2$), there is little
information about salinity in the temperature data, and therefore less
uncertainty reduction. In an application with fresh cold water from a
river source, the temperature and salinity variables will not only be
interdependent, but will also likely show dependence in the spatial
dimension. This in turn will impact the design criteria when we
evaluate the information measure by integrating over several locations
(Section \ref{sec:simulations}).

\subsubsection{Including Spatiality}

We now turn to an example involving a full-fledged gaussian random
field (GRF). The class of GRF model we will consider will generally
have a linear trend

\begin{align*}
\mu(s)=\mathbb{E}\left[\begin{pmatrix}
Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc
\end{align*}
with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we will only consider covariance models of separable type
\begin{align*}
\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
   \gamma_0 \sigma_i \sigma_j,~i\neq j
        \end{cases}
\end{align*}
where $k(., .)$ is one of the traditional spatial covariance kernel
(exponential, Mat\'{e}rn, ...) and $\gamma$ defines the
cross-covariance structure.

In the accompanying \kc{Where? On Github? If so say so.} Python
examples taking place within the MESLAS toolbox (URL \kc{Note}), these
modeling assumptions can be generalized to anisotropic covariance and
changing variance levels across the spatial domain. Anisotropy and
non-stationary variance are both relevant for the setting with river
plumes, but in practice this requires more parameters to be
specified. With extensive data and prior knowledge, one could also
possibly fit and estimate parameters of more complex multivariate
spatial covariance functions
\citep{gneiting2010matern,genton2015cross}, but that is outside the
scope of the current paper.

In the rest of this section, we will consider a GRF with mean and
covariance structure as above and parameters
\begin{align*}
\beta_0 = \begin{pmatrix}
5.8\\ 24.0
\end{pmatrix}, ~ \beta_1 = \begin{pmatrix}
0.0 & -4.0\\
0.0 & -3.8
\end{pmatrix},~ \sigma_1 = \sigma_2 = 2.25, ~ \gamma_0 = 0.2
\end{align*}
and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
One realization of this GRF is shown in Fig. \ref{example_excu}.

We now study how the expected Bernoulli variance eq.\eqref{def:eibv}
associated with data collection at a point changes if only one of the
two components of the field is observed. We first draw a realization
of the GRF defined above and use it as ground-truth to mimick the real
data-collection process. A first set of observations are done at the
locations depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data
is used to update the GRF model. We then consider the green triangle
as a potential next observation location and plot the expected
Bernoulli variance reduction (at each grid point \kc{I don't recall we
  have been explicit about grids before this mention. It might be
  useful to recall that here.}) that would result from observing only
one component of the field (temperature or salinity), or both at that
point.

\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_3.png}
  \caption{Pointwise Bernoulli variance reduction for observation of a
    single or both components of the random field at one
    location. Data collection locations in green. True excursion set
    in red. Places where only one response is above threshold are
    depicted in pink}
\label{fig:ebv_comp}
\end{figure}

Note that plotting the Bernoulli variance reduction at each point
might also be used to compare different data collection plans. For
example, Fig. \ref{fig:ebv_north_vs_east} shows the expected Bernoulli
variance reduction associated with a data collection plan along a
vertical line (static north) and one associated with a horizontal
(static east). Both expectations are computed according to the
a-priori distribution of the GRF (i.e. no observations have been
included yet).

\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east_2.png}
  \caption{Pointwise Bernoulli variance reduction for two different
    observation plans. Data collection locations in green. True
    excursion set is in red. Places where only one response is above
    threshold are depicted in pink.}
\label{fig:ebv_north_vs_east}
\end{figure}

In all the computed examples, the spatial domain $\domain$ has been
discretized to a set of $N$ grid locations
$\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell has area
$\delta$; the same grid is used for the waypoint graph for possible
design locations. The EIBV is approximated by sums over all grid
cells. \kc{This should be promoted up a couple of paras and would
  address my last comment above.}

\section{Sequential designs and heuristic path planning}\label{sec:heuristics}

The goal of this section is to present sequential data collection
strategies that aim at reducing the expected uncertainty on the target
ES $\es$.

\subsection{Background}
%As before, let $\gp$ be a $\no$-variate random field on $\domain$. See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.

From a sequential point of view, $n$ data collection steps have
already been performed and one wants to choose what data to collect
next. After each step, the GRF model is updated using co-Kriging
equations \ref{eq:meanCoK} and \ref{eq:varCoK}. The design evaluations
are based on the conditional expectation $\currentExp{.}$ from the law
of the field, conditional on the data gathered up to (and including)
stage $n$.

Note that the type of data collected at each stage can be of various
type (all components of the field at a single location, only some
components at a subset of the selected locations, etc.) since the
concept of \textit{generalized location} allows the co-Kriging
equations to handle all those case by putting them on equal footing.
In general, a design strategy must choose the spatial location as well
as the components to observe (heterotopic), or where several
observations are allowed at each stage (batch).  For the case with an
AUV exploring the river plume, we limit our scope to choosing one of
the neighboring spatial location (waypoints) at each stage, and all
components (temperature and salinity) of the field are observed there
(isotopic). The candidate points at this stage are denoted
$\candidates$ as defined from the 6 directions (apart from edges) in
the waypoint graph (see Fig.~\ref{fig:wp_graph_a} for an example). The
set $\candidates$ depends on the current location, but for readability
we suppress this in the notation.

The mathematical expression for the optimal design in this sequential
setting involves a series of intermixed maximizations over designs and
integrals over data. In practice, the optimal solution is intractable
because of the enormous growth over stages (see
e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
Instead, we outline heuristic strategies.

\subsection{A Naive Sampling Strategy}
\label{naive}

A simple heuristic for adaptive sampling is to observe $Z$ at the
location in $\candidates$ with current EP closest to
$\frac{1}{2}$. While easy to implement, this strategy can lead to
spending many stages in boundary regions regardless of the possible
effect of sampling at the considered point for the future conditional
distribution of $Z$. The strategy does not account for the expected
reduction in uncertainty, and it does not consider having an
integrated effect over other locations.

\subsection{Myopic Path Planning}
\label{sec:myopic}

The myopic (greedy) strategy which we present here is optimal if we
imagine taking only one more stage of measurements; it does not
anticipate what the subsequent designs might offer beyond the first
stage.  Based on the currently available data the myopic strategy
selects the location that leads to the biggest reduction in EIBV:
\begin{criterion}[Myopic]
The next observation location $\spatloc_{\stage + 1}$ is chosen among
the minimizers in $\candidates$ of the criterion: 
\begin{equation}\label{critSEQ}
     C_{\text{myopic}}(u) = \EIBV_{\stage}\left(\spatloc\right)
\end{equation}
\end{criterion}

The EIBV is efficiently computed for each of the candidate points
$\candidates$ using Proposition \ref{propo2}. Once the best location
has been selected, a stage of observation is performed and the GRF
model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$
after which the process is repeated.

Even though this myopic strategy is non-anticipatory, it still
provides a reasonable approach for creating designs in many
applications. Moreover, it can be implemented without any demands on
computational power, making it well-suited for embedding on an AUV.

\subsection{Look-ahead Trajectory Planning}\label{sec:LA}

We now extend the myopic strategy by considering two stages of
measurements, which is optimal in that it accounts consistently for
the expectations and minimizations in these two stages, but
nevertheless without including any planning beyond those two steps.

Assume data have been collected up to stage $n$. The principle of
two-step lookahead is to select the next observation location
$\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
we were to (optimally) add one more observation after the one at
$\spatloc_{\stage + 1}$. In order to formalize this concept, we must
extend the notation for EIBV in the future (after observation
$\stage + 1$ has been made). We let $\currentEIBV(\cdot; u, y)$ to
denote the EIBV where expectations are taken conditional on the data
available at $n$ and on an additional observation $y$ at $u$.

\begin{criterion}[2-step lookahead]
      The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
      \begin{align}\label{critLA}
          C_{\text{2-steps}}(u) &= \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
                  \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,
      Y\right)\right]
      \end{align}
      where $Y$ is a random variable distributed according to the
      conditional law of $\gp_{\spatloc}$ at step $\stage$ with the
      dependence of the set of candidates on the current location
      having been made explicit for the second stage of measurements.
    \end{criterion}

    In a practical setting, the first expectation can be computed by
    Monte Carlo sampling of data $Y$ from its conditional
    distribution. For each of these data samples, the second
    expectation is solved using the closed-form expressions for EIBV
    provided by Proposition \ref{propo2}, now with conditioning on the
    first stage data already going into the co-Kriging updating
    equations.

\subsection{Simulation study}
\label{sec:simulations}

\subsubsection{Static and Sequential Sampling Designs}
\label{sec:sampling_designs}

%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
%\end{figure}
%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.

We compare three different static designs denoted
\textit{static\_north}, \textit{static\_east}, and
\textit{static\_zigzag} with the three described sequential approaches
\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
AUV sampling paths are pre-scripted and cannot be altered.
For a fixed survey length, a closed-form expression for the EIBV is
available as in Proposition \ref{propo_eibv}. However, for the sequential
approaches this is not the case. For comparison the properties are
therefore evaluated using Monte Carlo sampling over several replicates
of realizations from the model while conducting simulated sequential
surveys for each one. We also compare predictive
performance measured by root mean square error (RMSE) for temperature
and salinity estimates as well as the variance reduction in these
two variables. It is important to note that the objective function
used by the robot. is focused on reducing the
EIBV, but we nevertheless expect that we will achieve good predictive
performance for criteria such as RMSE as well. Another non-statistical
criterion that is important for practical purposes is the computational
time needed for the strategy, as this will impact the performance for
an embedded system.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
%\caption{The equilateral waypoint graph used to discretize the
%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
%\label{fig:wp_graph}
%\end{figure}

\begin{figure}[!b] 
\centering 
\subfigure[The waypoint graph.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
\hfill
\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
\label{fig:wp_graph}
\end{figure}

Each strategy is conducted on an equilateral grid as shown in
Fig. \ref{fig:wp_graph}. The AUV starts at the center East-West
coordinate at the southern end of the domain (node 53). It then moves
along edges in the waypoint graph while collecting data which are
assimilated into the onboard GRF model. This is used in the evaluation
of the next node to sample.  The procedure is run for $10$ stages. A
total of 100 replicate simulations were conducted with all strategies.

\subsection{Simulation Results}

The results of the replicate runs are shown in
Fig. \ref{fig:sim_results}, where the different criteria are plotted
as a function of survey distance. Fig. \ref{fig:avg_ev} shows the
resulting drop in realized IBV for each of the six strategies. IBV
reduction occurs most under the \textit{myopic} and
\textit{look-ahead} strategies, each performing almost equally; this
is expected as the two criteria (Eq. \eqref{critSEQ} and
\eqref{critLA}) are sensitive to differences in IBV. The
\textit{static\_north} design also does well here because the path is
parallel to the boundary between the water masses.

\begin{figure}[!h]
  \centering
  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \hfill \subfigure[RMSE between estimated field and
  truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
  \hfill \subfigure[Explained variance
  $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
  \hfill \subfigure[Computational time for inference (some lines
  overlap \kc{Which ones? As figure with 6 keys, one sees only 4. This
    needs to be fixed or explained -- this was a repeat from an
    earlier
    feedback.}).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
\caption{Simulation results from 100 replicate simulations for 10
  sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
\label{fig:sim_results}
\end{figure}

Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
in RMSE and increase in explained variance, respectively. Both
\textit{myopic} and \textit{look-ahead} strategies perform well here,
but some of the \textit{static\_east} and \textit{static\_zigzag} also
achieve good results because they are cover large
parts of the domain without re-visitation. Sequential strategies
targeting IBV will sometimes not reach similar coverage, as
interesting data may draw the AUV path into twists and turns. There is
relatively large variety in the replicate results as indicated by the
vertical lines. Nevertheless, the ordering of strategies is similar.

Fig. \ref{fig:avg_time} shows the computational effort: the
\textit{naive} strategy is on par with the static designs, while the
\textit{myopic} strategy is slower because it evaluates expected values for all candidate directions at the waypoints. But it is still able to do so in reasonable time, which enables real-world applicability. The \textit{look-ahead} is even
slower, reaching levels that are nearly impractical for execution on a
vehicle. Some pruning of the graph is performed to improve the
performance, such as ruling out repeated visitations. Further pruning of branches or
inclusion of other heuristics could be included for better
performance. Then again, the inclusion of such heuristics is likely a
contributing factor for the \textit{look-ahead} strategy failing to
outperform the \textit{myopic} strategy.

%In Fig. \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown. The
%\textit{naive} strategy often gets stuck in the southern part of the
%domain because it is too focused on the probabilities near $0.5$. The
%\textit{myopic} strategy covers a wider domain than the naive or
%look-ahead. There are several reasons for this. First, a greedy
%approach will tend to put more emphasis on promising locations close
%to the agent, which may lead away from the centre. Second, as the
%agent evaluates the impact of locations further away (look-ahead)
%where assimilated data has less predictive power, the GP model (which
%is centered here) will act to restrict paths deviating from the
%central zone.

We studied the sensitivity of the results by modifying the input
parameters to have different correlations between temperature and
salinity, standard deviations, and spatial correlation range.  In all
runs, the \textit{myopic} and \textit{look-ahead} strategies perform
the best in terms of realized IBV, and much better than
\textit{naive}. The \textit{look-ahead} strategy seems to be
substantially better than the \textit{myopic} design only for very
small initial standard deviation or very large spatial correlation
range. 
%\textit{static\_north} continues to be the best static design
%for IBV, while \textit{static\_zigzag} is the best design for the
%other predictive performance measures, especially so with large
%spatial correlation range. 
We also ran simulation studies with only
temperature data, and for realistic correlation levels between
temperature and salinity, the IBV results are not much worse when only
temperature data are available. In addition to the comparison made in
Table \ref{tab:sim_rhoab}, the current setting includes spatial
correlation and this likely reduces the added effect of having
bivariate data. However, it seems that having temperature data alone
does a substantially worse job in terms of explained variance.

\section{Case Study - Mapping a River Plume}
\label{sec:case_study}

To demonstrate the applicability of using multivariate EPs and the IBV
to inform oceanographic sampling, we present a case study mapping a
river plume with an AUV. The experiment was performed in Trondheim,
Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
experiments were conducted in late Spring 2019, when there is still
snow melting in the surrounding mountains so that the river water is
substantially colder than the water in the fjord. The experiment was
focused along the frontal zone that runs more or less parallel to the
eastern shore as noted in Fig. \ref{fig:nidelven}.

\subsection{Model Specification}
\label{sec:exp_modeling}

The statistical model parameters were specified based on a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions and correlation
structures. Based on the initial data, the mean was specified by linear regression, where both temperature and salinity
were assumed to increase with the west location. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.

%\begin{figure}[!h] 
% \centering 
%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
%\caption{Data analysis from a preliminary trial experiment using the
%  AUV. Left: Residual plot of temperature and salinity along with
%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
%  distribution with two degrees of freedom. Right: Empirical variogram
%  of the salinity and temperature data.} \label{fig:parest}
%\end{figure}

\begin{figure}[!h]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
  \hfill
  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
  \caption{Data analysis from a preliminary trial experiment using the
    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Fig. \ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the westerly trends in both salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates larger variability in salinity than
temperature, and a positive correlation ($0.5$) between the two
variables. Based on the fitted bivariate Gaussian model (ellipses in
Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
the empirical CDF of these
quadratic forms (solid) together with the theoretical CDF of the
$\chi^2_2$ distribution. The modeled and theoretical curves are very
similar, which indicates that the Gaussian model fits reasonably
well. Fig. \ref{fig:parest_c} shows the empirical variogram of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m.
The working assumption of a separable covariance function is hence not unreasonable.

Based on the analysis in Fig. \ref{fig:parest}, the resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean. The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water. Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.

\begin{table}[!h]
\centering
\begin{tabular}{lrr}
\toprule
Parameter & Value & Source\\
\midrule
\rowcolor{Gray}
Cross correlation temperature and salinity & 0.5 & AUV observations\\
Temperature variance &  0.20 & AUV observations (variogram)\\
\rowcolor{Gray}
Salinity variance &  5.76 & AUV observations (variogram)\\
Correlation range  & 0.15 km & AUV observations (variogram)\\
\rowcolor{Gray}
River temperature  & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
Ocean temperature $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
\rowcolor{Gray}
River salinity $S_{river}$ & $14.0$ g/kg & AUV observations\\
Ocean salinity $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
\rowcolor{Gray}
Threshold in temperature & $10.5\,^{\circ}\mathrm{C}$ & User specified \\
Threshold in salinity & $18.0$ g/kg & User specified \\
\rowcolor{Gray}
\bottomrule
\end{tabular}
\caption{Model and threshold parameters from an initial AUV
  survey. Observations were taken across the front while crossing from
  fresh, cold river water to saline and warmer ocean waters.}
\label{tab:experiment_param}
\end{table}

\subsection{Experimental Setup}

A Light AUV \citep{sousa2012lauv}
(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
conductivity, temperature, and depth (CTD) sensor was used to provide
salinity and temperature measurements. We assume that the measurements are conditionally independent because the salinity is extracted from the conductivity instrument which is different from the temperature instrument. We specify variance $0.25^2$ for both errors, which is based on a middle ground between the nugget effect in the empirical variogram and the sensor specifications.

The AUV is a powered untethered platform that
operates at $1$-$3$ m/s in the upper water column. The AUV uses a multicore GPU NVIDIA
Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
The sampling algorithm was built on top of the autonomous agent framework
Teleo-Reactive EXecutive (\textit{T-REX})
\citep{py10,Rajan12,Rajan12b}.

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater
  Vehicle (LAUV) platform for upper water-column exploration used in
  our experiments.}
\label{fig:lauv}
\end{figure}

The AUV was running a
\textit{myopic} strategy to decide between sampling locations on the waypoint graph distributed over an equilateral grid, as shown in the grey-colored lattice in Fig. \ref{fig:map}. 
At each stage, it takes the AUV about 30
seconds to evaluate the EIBV for all the possible waypoint
alternatives.
The AUV was set to start in the south-center part of the waypoint
graph. A survey was set to take
approximately 40 minutes, visiting 15 waypoints on the grid, with the
AUV running near the surface to capture the plume. On its path from
one waypoint to the next, the AUV collected data with an update
frequency of 30 seconds, giving three measurements per batch in the updates at each stage.

\begin{figure*}[!h]
\centering
\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
\hspace{0.3cm}
\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}

%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
\caption{Results from mapping the Nidelva river, Trondheim, Norway
  over two survey missions. \ref{fig:map} shows an overview of the
  survey area overlaid with the AUV path in black and dashed
  line. Note the shaded region indicating a typical frontal
  region. \ref{fig:res_both} shows the collected temperature data as
  colored trails. Note waypoint 5 (WP5) which indicates where the two
  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
  collected salinity data overlaid on the final EP, which indicate the
  AUVs statistical impression of the front. For both missions the
  temperature and salinity data correspond with an indication of the
  EP front. About 2 hours time separated the two runs.}
\label{fig:results}
\end{figure*}

\subsection{Results}

Two survey missions (1 and 2), were run successively, with a short break in between. The resulting path of the
selected waypoints are shown in the map in Fig. \ref{fig:map}, both
within the expected frontal region (shaded pink). The recorded
temperatures are shown as colored trails in Fig. \ref{fig:res_both},
clearly indicating the temperature difference between fjord and
riverine waters. The salinity data are then shown separately, overlaid
with the estimated EP for each survey in Fig. \ref{fig:res1} and
\ref{fig:res2}.

Both surveys successfully estimated and navigated the separation zone,
crossing the frontal boundary multiple times. As conditions changed
slightly between the two surveys, the resulting path (after waypoint
5) is shown to deviate. Survey 1 continued northwards, tracking the
north-eastern portion of the front, while Survey 2 turned west,
mapping the south-western region.

The final predictions of the front location, represented by
conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
lines, correspond with one another. In both surveys they yield a
picture of the front being to the west in the southern portions of the
region and gradually bending off toward the north east. The amount of
exploration done by Survey 1 which turned north is greater than Survey 2 which was coming close to the survey area borders in the south-western
corner.

\section{Closing remarks}\label{sec:concl_disc}

This work builds on a multidisciplinary effort combining statistical
methods with robotic surveying techniques for oceanographic
applications. We show how observation practices can gain efficiency
and accuracy from the development of statistical techniques for spatial monitoring. We further demonstrate the
opportunities available for real-time multivariable spatial
data gathering and analysis onboard autonomous platforms, which
statisticians can exploit to create general-purpose toolkits for
similar applications.

In particular, we derive and show results for characterizing phenomena
connected to the properties of water masses. The characterization of
uncertainties in random sets is extended with new results for the expected integrated Bernoulli variance
reduction achieved by spatial sampling designs. This is
provided in closed-form for the situation with a static design, and
then extended to the adaptive situation. The sequential derivations
provide new insights into efficient applications of adaptive data collection,
as demonstrated in our application.

The case study consider the upper water column in the river plume, represented by a two dimensional grid. Extensions to three-dimensional domains are not methodologically different, but the operation must likely approximate calculations by integrating terms only in the vicinity of the autonomous vehicle \citep{fossum18b}. We did not consider any temporal effects, which would be
relevant on a larger time scale. We consider the extension to
spatio-temporal modeling as future work, and envision that
advection-diffusion equations could be useful in this kind of modeling
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex oceanographic
phenomena, the methods will need to be extended to non-Gaussian
phenomena, possibly feature-based mixtures of Gaussian processes which could still be
run onboard and augmented by dynamical models.

The spatial-statistical design criterion building on random sets is
relevant in our setting with different water properties. 
We show
mathematical generality beyond the expected integrated Bernoulli variance, for instance that of volume uncertainties which is possibly more relevant, but requires more computational resources.
Such criteria could be particularly useful in other oceanographic settings of algal-blooms, anoxic
zones or open water fronts \cite{costa19}.
Of course, other criteria
are also relevant for instance, hybrid or multi-attribute criteria
that could balance goals of exploration and exploitation in this
situation. Equally, such techniques have significant use cases in
downstream decision-making, with policy makers and regulators who need
to make difficult decisions related to fish farming or other marine
resource operations, and value of information analysis
\citep{Eidsvik:15} could be used to evaluate whether
information is likely to result in improved decision-making.
We also foresee opportunities related to design of experiments for multivariate processes using our notion of generalized locations. For instance, this is likely to be useful for internet-of-things applications or computer experiments context, where some observations or evaluations are rather inexpensive, while others are much more costly and must only be done when they are really valuable.

\kc{How about talking
  about removing any grids or waypoints altogether?}

In our context the myopic strategy perform rather well, and for computational reasons we did not go in depth on the dynamic program solution. There has been much creative work on finite horizon optimization in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling}, but their statistical properties are not yet clear.
It is equally interesting to
explore the additional flexibility that can be gained by having
multiple vehicles co-temporally exploring a spatial or spatio-temporal
domain \citep{ferreira2019advancing}. Such an approach would enable
concurrent sampling in different parts of the space, or opportunities
to move in parallel to best capture the excursion set.
The sometimes conflicting topics of autonomy and communication have caught much interest in control engineering lately \citep{zolich2019survey}, and statistical evaluations could contribute in the future. In our context, autonomy is highlighted, and this is reflected in our modeling and computational assumptions. With the possibilities of wifi communication to a mother-ship at waypoints, more advanced modeling and computational routines are of course possible, but with new challenges related to design and planning for when and where the communication should occur.

\section*{Acknowledgements}

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems
(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
Excellence, project number 223254, and the Applied Underwater Robotics
Labortatory (AURLab). JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. CT and DG acknowledge
support from the Swiss National Science Foundation, project number
178858.

%\begin{supplement}
%\sname{Supplement A}\label{suppA}
%\stitle{Title of the Supplement A}
%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
%\sdescription{Dum esset rex in
%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
%\end{supplement}

% == Adding references
\footnotesize
\bibliographystyle{imsart-nameyear}
\bibliography{ref}

\section*{Appendix}
\input{appendix}

% AOS,AOAS: If there are supplements please fill:
%\begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{Some text}
%\end{supplement}

% === Not used
% After reviews: - This is moved to Section 2.1.

%Traditional data collection at sea has typically been based on static
%buoys, Lagrangian floats, or ship-based methods, with significant
%logistical limitations that directly impact coverage and sampling
%resolution. Modern methods using satellite remote-sensing provide
%large-scale coverage but have limited resolution, are limited to
%sensing the surface of the ocean, and are impacted by cloud cover. The
%advent of robust mobile robotic platforms \citep{Bellingham07} has
%resulted in significant contributions to environmental monitoring and
%sampling. In particular, autonomous underwater vehicles (AUVs), have
%advanced the state of sampling and consequently have made robotics an
%integral part of ocean observation; our previous work has contributed
%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
%focusing on hierarchical statistical models; \cite{sahu2008space},
%studying spatio-temporal models for sea surface temperature and
%salinity data; and \cite{mellucci2018oceanic} looking at the
%statistical prediction of features using an underwater glider.

%\begin{itemize}
%
%\item Full numerical ocean models cannot provide accurate results
%  online if run on robotic sensing platforms, as onboard computers
%  cannot deliver the computational power required. Hence statistical
%  proxy models of the environment must be used for learning where to sample.
%
%\item With limited available information about the state of the ocean, there is substantial value in reacting to
%  information obtained from measurements taken in-situ. This
%  acquired information must be assimilated into statistical models that
%  can be used to inform decisions on where to sample
%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
%
%\item For sampling problems related to environmental sensing, the
%  number of choices (i.e. locations, trajectories, and candidate
%  designs) is enormous, creating a
%  trade-off between optimization (finding the most resource-efficient
%  design to collect necessary data) and computability (arriving at a
%  solution in reasonable time). To successfully resolve features, this
%  trade-off has to be considered in development and practice.
%
%\end{itemize}

%Addressing this, the combination of statistical tools and robotic platforms is a
%natural symbiosis which enables information-based sensing. Central to
%this is the ability to model spatially-correlated variables and
%provide formal measures of uncertainty. Our formulation is based on
%Gaussian Processes (GPs) as they allow efficient implementation 
%and evaluation in real time onboard a robotic platform.

%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.

%While the focus has often been on
%biological and anthropogenic impact from micro-plastics to
%pollution, biological oceanographers have focused intently on studying
%micro-organisms at the base of the human food web. These organisms are
%critically impacted by the changing dynamics in the upper water-column,
%especially in coastal zones which are complex and often hard to observe
%in space and time. By studying the bio-geochemical processes in the
%upper water-column scientists can measure the impact of change, natural
%or anthropomorphic, and provide an informed opinion to policy makers to
%effect changes in preserving the environment. However, the challenge of 
% The pressure on marine resources is growing and increased accuracy,
% resolution, and persistent monitoring of the oceans is crucial for
% long-term sustainable management.

% A
% sustained focus on prioritized and efficient data collection strategies
% have therefore started to emerge. The advent of marine robotic
% platforms, especially

% provided means to execute this prioritization through the capacity of
% autonomy and data-driven sampling, where data collection in principle
% can be optimized. These capabilities have made

% of the emerging sensing practice for ocean science, allowing scientists
% to increase the observational efficiency and resolution beyond what was
% previously possible. But how should a robotic platform, such as an AUV,
% effectively prioritize and identify important regions for sampling? The
% answer to this question relates to

% , and
% the application domain is clearly an arena where statisticians can
% contribute.
% There has recently been some statistical attention in oceanography:

%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest. Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity. One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}. These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively. Quantification of gradient features is therefore a much needed competence in robotic sampling.
\end{document}