### 2020-06-26 main_new.tex
Updates from Overleaf

\documentclass[aoas]{imsart}

\input{commands.tex}
\usepackage{amsmath}
\usepackage{pstricks,pst-grad}
\usepackage{graphicx}
\usepackage{floatrow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\floatsetup[table]{capposition=top}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}                     % horizontal lines in tables
\usepackage{comment}

% == Enable text degreehttps://www.overleaf.com/project/5a3268379ecbdc657d8767e8
\usepackage{textcomp}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% == Trygve Test
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

% put your definitions there:
\startlocaldefs
\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
\endlocaldefs

% New Marcos.
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cleveref}
\input{macros}

\begin{document}

\begin{frontmatter}

% "New Title of the paper"
\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields} 
\runtitle{Autonomous Oceanographic Data Collection}

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}

\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e2}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e4}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e5}}

\runauthor{TO. Fossum et al.}
\end{aug}

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for marine science and maritime resource management. Faced with
  limited resources in understanding processes in the water-column,
  the combination of statistics and autonomous systems provide new
  opportunities for experimental design. 
  %
  %In this work we develop methods for efficient spatial sampling applied to the mapping of coastal ocean processes by providing informative descriptions of spatial characteristics of ocean phenomena. 
  %
  % REPLACED BY 
  %
  In this work we develop efficient spatial sampling methods for characterizing regions defined by simultaneous exceedances above prescribed thresholds of several responses, with an application focus on mapping coastal ocean phenomena based on temperature and salinity measurements. 
  % EN REPLACEMENT
  Specifically, we define a design criterion based on uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected integrated Bernoulli variance reduction in such a framework. We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective. 
  We use simulations to study and compare properties of the considered approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a study mapping the boundary of a river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.

%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions. These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.

\end{abstract}

\begin{keyword}
\kwd{Ocean Sampling}
\kwd{Excursion Sets}
\kwd{Gaussian Processes}
\kwd{Experimental Design}
\kwd{Autonomous robots}
\kwd{Adaptive Information Gathering}
\end{keyword}

\end{frontmatter}
\section{Introduction}

%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.

Motivated by the challenges related to efficient data collection
strategies for our vast oceans, we combine spatial statistics, design
of experiments and marine robotics in this work.  The
multidisciplinary efforts enable information-driven data collection in
regions of high-interest.

\subsection{Oceanic data collection and spatial design of experiments}

Monitoring the world's oceans has gained increased importance in light
of the changing climate and increasing anthropogenic impact. Central
to understanding the changes taking place in the upper water-column is
knowledge of the bio-geophysical interaction driven by an
agglomeration of physical forcings (e.g. wind, topography, bathymetry,
tidal influences, etc.) and incipient micro-biology driven by
planktonic and coastal anthropogenic input, such as pollution and
agricultural runoff transported into the ocean by rivers.  These often
result in a range of ecosystem-related phenomena such as blooms and
plumes, with direct and indirect effects on society \citep{ryan2017}. One of the
bottlenecks in the study of such phenomena lies however in the lack of
observational data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations in which ocean processes transpire, prompting the need for
effective means of data collection.  By \emph{sampling}, we refer here
primarily to the design of observational strategies in the spatial
domain with the aim to pursue measurements with high scientific
relevance.  Models and methods from spatial statistics and experimental design can clearly contribute to this sampling challenge.

%applications and beyond.  
 %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance. This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.

Data collection at sea has typically been based on static buoys,
floats, or ship-based methods, with significant logistical limitations
that directly impact coverage and sampling resolution. Modern methods
using satellite remote-sensing provide large-scale coverage but have
limited resolution, are limited to sensing the surface, and are
impacted by cloud cover. Numerical ocean models similarly find it
challenging to provide detail at fine scale \citep{Lermusiaux:2006},
and also come with computational costs that can be limiting. The
advent of robust mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling in the ocean (Fig.~\ref{fig:envir1}). In particular,
autonomous underwater vehicles (AUVs) have advanced the state of data
collection and consequently have made robotics an integral part of
ocean observation \citep{das11b,Das2015,fossuminformation,fossum18b}.

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of a range of ocean sensing opportunities.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on 
    ship-based sampling has been augmented by autonomous
    robotic vehicles such as AUVs. % and their interactions.  
    %AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
    % using the sense-plan-act control approach to autonomous control. 
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective oceanographic observation.}
  \label{fig:envir} \end{figure}

Surveys with AUVs are usually limited to observations along fixed
transects that are pre-scripted in mission plans created manually by a
human operator. Missions can be specified operating on a scale of
hundreds of meters to tens of kilometers depending on the scientific
context. Faced with limited coverage capacity, a more effective
approach is to instead use onboard algorithms to continuously
evaluate, update, and refine future sampling locations, making the
data collection \emph{adaptive}.  In doing so, the space of sampling
opportunities is still limited by a waypoint graph, which forms a
discretization of the search domain where the AUV can navigate;
however the AUV can now modify its path at each waypoint based on
in-situ measurements and calculations onboard using onboard
deliberation \citep{py10,Rajan12,Rajan12b}.  Full numerical ocean
models based on complex differential equations cannot be run onboard
the AUV with limited computational capacity, and statistical models
relying on random field assumptions are relevant as a means to
effectively update the onboard model from in-situ data, and to guide
AUV data collection trajectories.

The work presented here is primarily inspired by a case study
pertaining to using an AUV for spatial characterization of a frontal
system generated by a river plume. Fig.~\ref{fig:nidelven} shows the
survey area in Trondheim, Norway, where cold freshwater enters the fjord from a river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coriolis force the
cold fresh water tends to flow east. Depending on the variations in
river discharge, tidal effects, coastal current and wind, this
boundary often gets distorted, and knowledge about its location is
highly uncertain, making deterministic planning challenging. The goal
is therefore to use AUV measurements for improved description of the
interface between fresh and oceanic waters.  It is often not possible
to sample the biological variables of fundamental interest in such AUV
operations, but off-the-shelf instruments provide temperature and
salinity measurements which serve as proxys for the underlying
biological phenomenon. With the help of a vector-valued random field
model for temperature and salinity, one can then aim to describe the
plume.  The goal of plume characterization, in this way, relates to
that of estimating some regions of the domain, typically excursion
sets (ESs), when implicitly defined by the vector-valued random field.
In our context of environmental sampling, the joint salinity and
temperature excursions of a river plume help characterize the
underlying bio-geochemical processes
\citep{hopkins2013detection,Pinto2018}. Motivating examples for ESs of
multivariate processes are also abundant in other contexts, for
instance in medicine, where physicians do not rely solely on a single
symptom but must see several combined effects before making a
diagnosis.

The questions tackled here hence pertain to the broader area of
spatial data collection for vector-valued random fields.
%Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs. 
Given the operational constraints on AUV movements and the fact that
surveys rely on successive measurements along a trajectory, addressing
corresponding design problems calls for sequential strategies.  Our
main research angle in the present work is to extend sequential design
strategies from the world of spatial statistics and computer
experiments to the setting of both vector-valued observational data
and experimental designs for feasible robotic trajectories. We
leverage and extend recent progress in expected uncertainty reduction
for ESs of Gaussian random fields (GRFs) in order to address this
research
problem. %approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
We briefly review recent advances in targeted sequential design of
experiments based on GRFs before detailing other literature related to
AUV sampling and our contributions prior to outlining the rest of the
paper.

\subsection{Random field modeling and targeted sequential design of experiments}

While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even
for vector-valued random field models with associated prediction
approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003},
there has lately been a renewed interest for random field models in
the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in
simulation experiments \citep{Santner.etal2003}. As detailed in
\cite{Ginsbourger2018}, GRF models have been used in particular as a
basis to sequential design of simulations dedicated to various goals
such as global optimization and set estimation. Of particular relevance to our context,
\cite{Bect.etal2012} focuses on strategies to reduce uncertainties related to volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} concentrates on making the latter strategies computationally efficient and batch-sequential. Rather than focusing on excursion volumes, approaches were investigated in
\cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
with ambitions of estimating sets themselves. Recently, sequential
designs of experiments for the conservative estimation of ESs based on
GRF models were presented in \citep{Azzimonti.etal}.

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, the PhD thesis \citep[][p.82]{stroh} mentions general
possibilities of stepwise uncertainty reduction strategies for ES's in
the context of designing fire simulations, yet outputs are mainly
assumed independent.

% From earlier version; to be relocated?
%Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian Random Field (GRF) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes. 
\subsection{Previous work in AUV sampling}

Other statistical work in the oceanographic domain include
\cite{wikle2013modern} focusing on hierarchical statistical models,
\cite{sahu2008space} studying spatio-temporal models for sea surface
temperature and salinity data and \cite{mellucci2018oceanic} looking
at the statistical prediction of features using an underwater glider.
In this paper the main focus is not on statistical modeling per se, but rather on statistical principles and computations underlying efficient data collection. We combine novel possibilities in marine
robotics with spatial statistics and experimental design to provide
useful AUV sampling designs.

Adaptive in-situ AUV sampling of an evolving frontal feature has been
explored in \cite{fronts11,Smith2016,Pinto2018,costa19}. These
approaches typically use a reactive-adaptive scheme, whereby
exploration does not rely on a statistical model of the environment,
but rather adaptation is based on closing the sensing and actuation
loop. Myopic sampling, i.e. stage-wise selection of the path (on the
waypoint graph), has been used for surveys
\citep{singh2009efficient,Binney2013} that focus largely on reducing
predictive variance or entropy. These criteria are widely adopted in
the statistics literature on spatio-temporal design as well
\citep{bueso1998state,zidek2019monitoring}. However, response variance and entropy being depending only in GRF models on measurement locations and not on response values, criteria based on them only tend to have limited flexibility for
active adaptation of trajectories based on measurement values.  The use
of data-driven adaptive criteria was introduced to include more
targeted sampling of regions of scientific interest in \cite{Low2009}
and \cite{fossuminformation}.

The primary contributions of this work are:

\begin{itemize}
\item Extending uncertainty reduction criteria to
  vector-valued cases.  
\item Closed-form expressions for the expected integrated Bernoulli
  variance (IBV) of the excursions in GRFs. 
\item Algorithms for myopic and multiple-step ahead sequential
  strategies for optimizing AUV sampling with respect to the mentioned
  criteria. 
\item Replicable experiments on synthetic cases with accompanying
  code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
\item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
\end{itemize}

The remainder of this paper is organized as follows:
Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and
the design criteria connected to the IBV for excursions of
vector-valued GRFs. Section \ref{sec:heuristics} builds on these
assumptions when deriving the sequential design criteria for adaptive
sampling. In both sections properties of the methods are studied using
simulations. Section \ref{sec:case_study} demonstrates the methodology
used in field work characterizing a river plume. Section
\ref{sec:concl_disc} contains a summary and a discussion of future
work.

\section{Quantifying uncertainty on Excursion Sets implicitly defined by GRFs}
\label{sec:ESEP}

Section \ref{sec:bg_and_notation} introduces notation and co-Kriging
equations of multivariate GRFs.  Section \ref{sec:set_uq} presents
uncertainty quantification (UQ) techniques on ESs of GRFs, in
particular the IBV and the excursion measure variance (EMV).  Section
\ref{sec:eibv} turns to the effect of new observations on EMV and IBV,
and semi-analytical expected EMV and IBV over these observations are
derived.
%The resulting formulae form the backbone of the expected sequenatil uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate
example relevant for temperature and salinity in our case.

\subsection{Background, Notation and co-Kriging}
\label{sec:bg_and_notation}

We denote by $\gp$ a vector-valued random field indexed by some
arbitrary domain $\domain$, and assume values of the field at any
fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a
$\no$-variate random vector ($\no\geq 2$). In the river plume
characterization case, $\domain$ is a prescribed domain in
Trondheimsfjord, Norway (for the purpose of our AUV application, a
discretization of a $2$-dimensional domain at fixed depth is
considered), and $\no=2$ with responses of temperature and salinity. A
bivariate GRF model is assumed for $\gp$. To motivate concepts,
Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} shows a realization of such a
vector-valued GRF on $\domain=[0,1]^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
realizations, namely regions: i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water, ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in light-red, where one variable is above and the other below their respective thresholds, indicative of mixed waters.

\begin{figure}[!b] 
\centering 
\subfigure[Temperature.]{\includegraphics[
height=0.27\textwidth,keepaspectratio]{Figures/intro_temp.png}\label{fig:real_temp}}
\subfigure[Salinity.]{\includegraphics[
height=0.27\textwidth,keepaspectratio]{Figures/intro_sal.png}\label{fig:real_sal}}
\subfigure[Regions of interest.]{\includegraphics[
height=0.27\textwidth,keepaspectratio]{Figures/intro_excu.png}\label{fig:jointex_roi}}
\caption{Realization of a bivariate GRF (display (a) and (b)) and excursion set above some threshold (c). Joint excursion in red and excursion of a single variable in light-red.}
\label{example_excu}
\end{figure}

%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
For the general setting of a $\no$-variate random field, we are
interested in recovering the set of locations $\es$ in the domain for
which the components of $\gp$ lie in some set of specified values
$\T\subset \mathbb{R}^{\no}$; in other words \textit{the pre-image of
  $T$ by $\gp$}:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
If we assume that $\gp$ has continuous trajectories 
and $T$ is closed, then $\es$ becomes a Random Closed Set
\citep{Molchanov2005} and concepts from the theory of random sets will
prove useful to study $\es$.
%under some measure defined on the domain.
%\textcolor{red}{Is this really needed?} 
Note that while some aspects of the developed approaches do not call
for a specific form of $\T$, we will often, for purposes of
simplicity, stay with the case of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of
$\gp[\spatloc]$ ($1\leq \ell\leq \no$), we use the term
\textit{generalized location} for the couple $x=(\spatloc,\ell)$.
%Given such a generalized location notation $x$,  
The notation $\gp[x]$ will be used to denote $\gp[\spatloc,\ell]$ 
and
%
%This slight change of notation 
will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the one of univariate Kriging. %Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$. In the following, the letter $x$ will be reserved for generalized locations, while 
The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
%
% Replacement by overlooked revised chunks from June 19th
%\begin{comment}
%Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
%\begin{align*}
%\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
%\end{align*}
%In general,  corresponding to batches of observations. %The field values of batch observations are denoted by
%\begin{align*}
%\gp[\bm{x}]:=
%\left(\gp[\spatloc_1,\ell_1], ...,
%\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
%\end{align*}
%In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
%Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance . This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
%We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
%Assuming that one wishes 
%to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
%given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
%The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
%
%\begin{equation}\label{eq:cokrig_mean}
%\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
%\end{equation}
%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
%where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations. 
%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
%The associated co-Kriging %(conditional) 
%residual cross-covariance function can be expressed in the same vein via
%
%\begin{equation}\label{eq:cokrig_cov}
%k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
%\end{equation}
%\subsubsection{co-Kriging update formulae}
%In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations. Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case. Because of our representation of co-Kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}. The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$. The mean update formulae is
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
% delivers that
% 
%\begin{equation}\label{eq:meanCoK}
%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
%\end{equation}
%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
%The updated co-Kriging residual cross-covariance function is
%\begin{eqnarray}\label{eq:varCoK}
%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
%\end{eqnarray}
%\medskip
%
%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
%\end{comment}
%
%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
%one for a single observation, provided some notation adjustments are made. This motivates the following.
Furthemore, boldface letters will be used to denote concatenated
quantities corresponding to batches of observations.  Given a dataset
consisting of
$q$ observations at spatial locations
$\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in
\domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in
\lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
we use the concatenated notation
\begin{align*}
\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
\end{align*}
We also compactly denote the field values at those different locations by
\begin{align*}
\gp[\bm{x}]:=
\left(\gp[\spatloc_1,\ell_1], ...,
\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
\end{align*}

For a second order random field $(Z_{\spatloc})_{\spatloc \in
  \domain}$ with mean $\mu$ and matrix covariance function $K$,
$\mu$ is naturally extended to $\domain \in \lbrace 1, ...,
\no\rbrace$ into a function of $x=(\spatloc,
\ell)$ and is further straightforwardly vectorized into a function of
$\bm{x}$. As for $K$, it induces a covariance kernel
$k$ on the set of extended locations via $k((\spatloc,
\ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell,
  \ell'}$. In vectorized/batch form, $k(\bm{x},
\bm{x}')$ then amounts to a matrix with numbers of lines and columns equal to
the numbers of generalized locations in
$\bm{x}$ and
$\bm{x}'$, respectively. Such
vectorized quantities turn out to be useful in order to arrive at
simple expressions for the co-Kriging equations below.
%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}

Given a GRF $\gp$ and observations of some of its components at locations in the domain, one can predict the value of the field at some
unobserved location $\spatloc\in \domain$ by using the conditional
mean of $\gp[\spatloc]$, conditional on the data. This coincides with co-Kriging equations,
which tell us precisely how to compute conditional means and covariances.  We will present a
general form of co-Kriging, in the sense that it allows inclusion of
several (batch) observations at a time; observations at a given
location $u \in \domain$ may only include a subset of the components
of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).

Assuming that $n$ batches of observations are available with sizes
$q_1,\dots, q_n$, and that one wishes to predict $\gp[\bm{x}]$ for
some batch of $q\geq 1$ generalized locations
$\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
the simple co-Kriging mean then amounts to Kriging with respect to a
scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
%
\begin{equation}\label{eq:cokrig_mean}
\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})).
\end{equation}
Here, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
q_i$)-dimensional vector of observed (noisy) responses of
$Z$ at all considered generalized locations, and
$\lambda_{[n]}(\bm{x})$ is a vector of weights equal to
$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
$$
with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
\bm{x}_n)$ and where
$\Delta_{[n]}$ is the covariance matrix of
Gaussian-distributed noise assumed to have affected measurements up to
batch
$n$. For our applications with salinity and temperature observations, this matrix is diagonal because we assume conditionally independent sensor readings, but it might not be diagonal with other types of combined measurements.
The matrix in parenthesis will be assumed to be non-singular
throughout the presentation. The associated co-Kriging %(conditional)
residual (cross-)covariance function can also be expressed in the same vein via
%
\begin{equation}\label{eq:cokrig_cov}
k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T 
\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)
\lambda_{{[n]}}(\bm{x}').
\end{equation}

%\subsubsection{co-Kriging update formulae}

Let us now consider the case where a co-Kriging prediction of $Z$ was
made with respect to $n$ batches of generalized locations, concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
and one wishes to update the prediction by incorporating a new vector
of observations $\mathbf{z}_{n+1}$ measured at a batch of
$q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
Thanks to our representation of co-Kriging in terms of simple Kriging
with respect to generalized locations, a strightforward adaptation of
the batch-sequential Kriging update formulae from
\citep{Chevalier.etal2013a} suggests that
% 
\begin{equation}\label{eq:meanCoK}
\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
\end{equation}
where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
to the Kriging weights for the last $q_{n+1}$ responses
when predicting at $\bm{x}$ relying on all measurements until batch
$(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
The associated co-Kriging residual (cross-)covariance function is
\begin{align}\label{eq:varCoK}
k_{[n+1]}(\bm{x},\bm{x}') & = k_{[n]}(\bm{x},\bm{x}')\\
 \nonumber - & \lambda_{[n+1,n+1]}(\bm{x})^T 
\left(k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}\right)
%k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
\lambda_{{[n+1,n+1]}}(\bm{x}'),
\end{align}
%\medskip
%
As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
these update formulae naturally extend to universal Kriging in
second-order settings and apply without Gaussian assumptions. We will
now see how the latter formulae are instrumental in deriving
semi-analytical formulae for step-wise uncertainty reduction criteria
for vector-valued random fields.

\subsection{Uncertainty Quantification on ESs of multivariate GRFs}
\label{sec:set_uq}

We now introduce quantities that allow UQ on the volume of the ES
$\es$. Let $\mes$ be a (locally finite, Borel) measure on
$\domain$. We want to investigate the probability distribution of
$\mes(\es)$ through its moments.  Centered moments may be computed
using Proposition~\ref{propo1} developed in the appendix.  In
particular, as an integral over EPs, the 
$\EMV = \operatorname{Var}[\mes(\es)]$ is:
\begin{equation*}
\begin{split}
\EMV
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ is
\begin{equation*}
\begin{split}
\EMV
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative
distribution function (CDF) numerically \citep{genz2009computation}.

Note that this quantity requires the solution of an integral over
$\domain^2$. In contrast, the IBV of
\cite{bect2019} involves solely an integral on $\domain$ and can be
expanded as
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%

\subsection{Expected IBV and EMV}
\label{sec:eibv}

We compute the expected effect of the inclusion of new observations on
the $\EMV$ and $\IBV$ of the ES $\es$. Let us consider the
same setting as in Eq. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
expectation and probability conditional on the first $n$ batches of
observations, respectively. %We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$. 
We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.

\medskip

In order to study the effect of the inclusion of a new data point, we
let $ \currentIBV(\bm{x}; \bm{y}) $ denote the expected IBV under the
current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$
(generalized, possibly batch observation). The expected effect of a
new observation on the IBV is then 
\begin{equation}\label{def:eibv}
    \currentEIBV(\bm{x}):=\currentExp{\mbox{IBV}(\bm{x}; \bm{Y})},
\end{equation}
where $\bm{Y}$ is distributed according to the current law of
$Z_{\bm{x}}$ and with independent noise having covariance matrix
$\Delta_n$.

We next present a result that allows efficient computation of $\EIBV$
as an integral of CDFs of the multivariate Gaussian distribution. This will prove
useful when designing sequential expected uncertainty reduction strategies.

\begin{propo}
\label{propo_eibv}
\begin{equation}
\begin{split}
\currentEIBV(\bm{x})
&=\int_{\domain} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
&-\int_{\domain} \varPhi_{2\no}
\left(
\left(
\begin{matrix}
\bt-\currentMean{u}\\
\bt-\currentMean{u}
\end{matrix}
\right);
\mathbf{\Sigma}_{[n]}(\uu)
\right)
d\mes(u),
\end{split}
\end{equation}
where the matrix $\mathbf{\Sigma}_{[n]}(\uu)$ is defined as
\begin{equation*}
\begin{split}
\mathbf{\Sigma}_{[n]}(\uu)&=
\left(
\begin{matrix}
\currentCov{u, u} & \currentCov{u, u}-\futureCov{u, u}\\
\currentCov{u, u}-\futureCov{u, u} & \currentCov{u, u}
\end{matrix}
%\begin{matrix}
%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
%\Delta_{[n]}(\uu) & \currentCov{u, u}
%\end{matrix}
\right).\\
\end{split}
\end{equation*}
\end{propo}

As for the expected EMV, a similar result may be derived.
\begin{propo}
\label{propo_emv}

\begin{equation*}
\begin{split}
\currentEEMV(\bm{x})
&=\int_{\domain^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); ~ \mu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\int_{\domain^2} \varPhi_{2\no}
\left(
\left(
\begin{matrix}
\bt-\currentMean{\uu}\\
\bt-\currentMean{\vv}
\end{matrix}
\right);~
\mathbf{\tilde{\Sigma}}_{[\stage]}(\uu, \vv)
\right)
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)
\end{split}
\end{equation*}
where the matrix $\mathbf{\tilde{\Sigma}}_{[n]}(\uu, \vv)$ is defined blockwise as
\begin{equation*}
\begin{split}
\mathbf{\tilde{\Sigma}}_{[n]}(\uu, \vv)&=
\left(
\begin{matrix}
\tilde{\Sigma}_{1,1}(\uu, \uu) & \tilde{\Sigma}_{1,2}(\uu, \vv)\\
\tilde{\Sigma}_{2,1}(\vv, \uu) & \tilde{\Sigma}_{2,2}(\vv, \vv)
\end{matrix}\right)
\end{split}
\end{equation*}
with blocks given by
\begin{equation*}
\begin{split}
\tilde{\Sigma}_{i,j}(u, v) &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv) + \delta_{i,j}\futureCov{\uu, \vv}
\end{split}
\end{equation*}
and the other blocks are defined by following the same logic, exchanging the role of $\uu$ and $\vv$.
\end{propo}

We remark that Propositions \ref{propo_eibv} and \ref{propo_emv} are
twofold generalizations of results from \cite{chevalier2014fast}: they
extend previous results to the multivariate setting and also allow for
the inclusion of batch or heterotopic observations through the concept
of generalized locations.  A key element for understanding these
propositions is that the conditional co-Kriging mean entering in the
EPs depend linearly on (batch) observations. The conditional equality
expressions thus become linear combinations of Gaussian variables
whose mean and covariance are easily calculated.  Related closed-form
solutions have been noted in similar contexts
\citep{bhattacharjya2013value,stroh}, but not generalized to our
situation with random sets for vector-valued GRFs.

\subsection{Expected Bernoulli variance for a two dimensional Example}
\label{Sec:UnivarEx}

We illustrate the expected Bernoulli variance (EBV) associated with different designs on a
bivariate example. This mimics our river plume application and hence
the first and second component of the random field will be called
\textit{temperature} and \textit{salinity}. We begin
with a \textit{pointwise} example, considering a single bivariate Gaussian distribution (i.e. no spatial elements).

\subsubsection{A pointwise study}

Say we want to study the excursion probability of a bivariate Gaussian
above some threshold, where the thresholds are set equal to the mean;
$\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
salinity, and we play with the temperature and salinity correlation
and variances to study the effect on the EP and EBV.

\begin{figure}[!b] 
\centering
\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
\caption{Density contour plots with increasing correlations between
  temperature and salinity. The densities have unit variance and
  thresholds identical to the mean values $5^o C$ and
  $30$ g/kg.}
\label{illus_bivarDens}
\end{figure}

Fig.~\ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity. The displayed densities have unit standard deviations for
both temperature and salinity, but we also study the effect of
doubling the standard deviations.

Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
Bernoulli variance (second row) for the examples indicated in
Fig.~\ref{illus_bivarDens}. The EPs increase with the correlation as
there is a strong tendency to have jointly low or high temperature and
salinity. The Bernoulli variance is similarly larger for high
correlations. EPs and Bernoulli variances are the same for temperature and salinity standard
deviations $\sigma_1$ and $\sigma_2$, which implies that high variability in
temperature and salinity is not captured in the $p(1-p)$ expression.

\begin{table}[!t] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and EBVs for both temperature and salinity data, and only temperature data (bottom rows).}
  \begin{tabular}{c|ccc|ccc}
 &\multicolumn{3}{c}{$\sigma_1=\sigma_2=1$} & \multicolumn{3}{c}{$\sigma_1=\sigma_2=2$} \\
\hline
Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
\hline
$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
EBV, Temperature and Salinity & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
EBV, Temperature only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
\hline
\end{tabular}
\label{tab:sim_rhoab}
\end{table}
The bottom two rows of Table \ref{tab:sim_rhoab} show 
EBV results. This is presented for a design gathering
both data types, and for a design with temperature measurements
alone. When both data are gathered, the measurement model is
$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with
$\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$,
$\epsilon \sim N(0,0.5^2)$ when only temperature is measured.  For
this illustration, Table \ref{tab:sim_rhoab} shows that the expected
Bernoulli variance gets smaller with larger standard deviations. The expected reduction of Bernoulli variance is
further largest for the cases with high correlation $\gamma$. Albeit smaller,
there is also uncertainty reduction when only temperature is measured
(bottom row), especially when temperature and salinity are highly
correlated. When correlation is low ($\gamma=0.2$), there is little
information about salinity in the temperature data, and therefore less
uncertainty reduction. 
%In an application with fresh cold water from a river source, the temperature and salinity variables will not only be interdependent, but will also likely show dependence in the spatial dimension. This in turn will impact the design criteria when we
% evaluate the information measure by integrating over several locations (Section \ref{sec:simulations}).

\subsubsection{Including Spatiality}

We now turn to an example involving a full-fledged GRF. The statistical model we consider has a linear trend
\begin{align*}
\mu(s)=\mathbb{E}\left[\begin{pmatrix}
Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc,
\end{align*}
with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we only consider separable covariance models;
\begin{align*}
\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
   \gamma \sigma_i \sigma_j,~i\neq j,
        \end{cases}
\end{align*}
where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is used, for Euclidean distance $h$.
In the accompanying Python
examples taking place within the MESLAS toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}, these
modeling assumptions can however be relaxed to anisotropic covariance and
changing variance levels across the spatial domain. Both extensions are relevant for the setting with river
plumes, but in practice this requires more parameters to be specified. With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
spatial covariance functions
\citep{gneiting2010matern,genton2015cross}, but that is outside the
scope of the current paper.

In the rest of this section, we consider a GRF with mean and
covariance structure as above and parameters
\begin{align*}
\beta_0 = \begin{pmatrix}
5.8\\ 24.0
\end{pmatrix}, ~ \beta_1 = \begin{pmatrix}
0.0 & -4.0\\
0.0 & -3.8
\end{pmatrix},~ \sigma_1 = 2.5,~ \sigma_2 = 2.25, ~ \gamma = 0.2,
\end{align*}
and kernel parameter $\eta=3.5$.
One realization of this GRF is shown in Fig.~\ref{example_excu}.
In the computed examples, the spatial domain $\domain$ is
discretized to a set of $N$ grid locations
$\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell has area
$\delta$; the same grid is used for the waypoint graph for possible
design locations. The EIBV is approximated by sums over all grid
cells.

We now study how the EBV [Eq.\eqref{def:eibv}]
associated with data collection at a point changes if only one of the
two components of the field is observed. We first draw a realization
of the GRF defined above and use it as ground-truth to mimic the real
data-collection process. A first set of observations are done at the
locations depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data
is used to update the GRF model. We then consider the green triangle
as a potential next observation location and plot the EBV reduction (at each grid node in the waypoint graph) that would result from observing only one component of the field (temperature or salinity), or both at that point.

\begin{figure}[!b] 
\centering 
\subfigure[Regions of interest.]{\includegraphics[
height=0.21\textwidth,keepaspectratio]{Figures/ebv_comp_excu.png}\label{fig:ebv_comp_excu}}
\subfigure[Temperature.]{\includegraphics[
height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_temp.png}\label{fig:ebv_comp_temp}}
\subfigure[Salinity.]{\includegraphics[
height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_sal.png}\label{fig:ebv_comp_sal}}
\subfigure[Both.]{\includegraphics[
height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_both.png}\label{fig:ebv_comp_both}}
\caption{Pointwise Bernoulli variance reduction for observation of a
    single or both components of the random field at one
    location. Data collection locations in green. True excursion set
    in red. Places where only one response is above threshold are
    depicted in pink. EBV reduction associated to observing one or both responses at the green location are shown in \ref{fig:ebv_comp_temp}, \ref{fig:ebv_comp_sal} and \ref{fig:ebv_comp_both}.}
\label{fig:ebv_comp}
\end{figure}

Note that plotting the EBV reduction at each point
might also be used to compare different data collection plans. For
example, Fig. \ref{fig:ebv_north_vs_east} shows the EBV reduction associated with a data collection plan along a vertical line (static north) and one associated with a horizontal
(static east). Both expectations are computed according to the
a-priori distribution of the GRF (i.e. no observations have been
included yet).

\begin{figure}[!b] 
\centering 
\subfigure[Excursion probability.]{\includegraphics[
height=0.25\textwidth,keepaspectratio]{Figures/ebv_static_excu.png}\label{fig:ebv_static_excu}}
\subfigure[Static north design.]{\includegraphics[
height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_north.png}\label{fig:ebv_static_north}}
\subfigure[Static east design.]{\includegraphics[
height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
  \caption{Pointwise Bernoulli variance reduction for two different observation plans. The prior EP is shown in \ref{fig:ebv_static_excu}. EBV reduction for each design shown in \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
\label{fig:ebv_north_vs_east}
\end{figure}

\section{Sequential designs and heuristic path planning}\label{sec:heuristics}

We present sequential data collection
strategies that aim at reducing the expected uncertainty on the target
ES $\es$.

\subsection{Background}
%As before, let $\gp$ be a $\no$-variate random field on $\domain$. See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.

From a sequential point of view, $n$ data collection steps have
already been performed and one wants to choose what data to collect
next. The design evaluations
are based on the conditional expectation $\currentExp{.}$ from the law
of the field, conditional on all data available at
stage $n$.
Once the best design at stage $n$
has been selected, the data are collected and the GRF
model is updated using co-Kriging
Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{\stage + 1}$
after which the process is repeated.

Note that the type of data collected at each stage can be of various
type (all components of the field at a single location, only some
components at a subset of selected locations, etc.) because of the
concept of \textit{generalized location} in the co-Kriging expressions.
In general, a design strategy must choose the spatial location as well
as the components to observe (heterotopic), or where several
observations are allowed at each stage (batch).  For the case with an
AUV exploring the river plume, we limit our scope to choosing one of
the neighboring spatial location (waypoints) at each stage, and all
components (temperature and salinity) of the field are observed
(isotopic). The candidate points at this stage are denoted
$\candidates$ as defined from the 6 directions (apart from edges) in
the waypoint graph (see Fig.~\ref{fig:wp_graph_a}). The
set $\candidates$ depends on the current location, but for readability
we suppress this in the notation.

The mathematical expression for the optimal design in this sequential
setting involves a series of intermixed maximizations over designs and
integrals over data. In practice, the optimal solution is intractable
because of the enormous growth over stages (see
e.g. \cite{powell2016perspectives}).
Instead, we outline heuristic strategies.

\subsection{A Naive Sampling Strategy}
\label{naive}

A simple heuristic for adaptive sampling is to observe $Z$ at the
location in $\candidates$ with current EP closest to
$\frac{1}{2}$. While easy to implement, this strategy can lead to
spending many stages in boundary regions regardless of the possible
effect of sampling at the considered point for the future conditional
distribution of $Z$. The strategy does not account for the expected
reduction in uncertainty, and it does not consider having an
integrated effect over other locations.

\subsection{Myopic Path Planning}
\label{sec:myopic}

The myopic (greedy) strategy which we present here is optimal if we
imagine taking only one more stage of measurements; it does not
anticipate what the subsequent designs might offer beyond the first
stage.  Based on the currently available data the myopic strategy
selects the location that leads to the biggest reduction in EIBV:
\begin{criterion}[Myopic]
The next observation location $\spatloc_{\stage + 1}$ is chosen among
the minimizers in $\candidates$ of the criterion: 
\begin{equation}\label{critSEQ}
     C_{\text{myopic}}(u) = \EIBV_{\stage}\left(\spatloc\right)
\end{equation}
\end{criterion}

The EIBV is efficiently computed for each of the candidate points
$\candidates$ using Proposition \ref{propo2}. 
Even though this myopic strategy is non-anticipatory, it still
provides a reasonable approach for creating designs in many
applications. Moreover, it can be implemented without too much demand on
computational power, making it well-suited for embedding on an AUV.

\subsection{Look-ahead Trajectory Planning}\label{sec:LA}

We now extend the myopic strategy by considering two stages of
measurements, which is optimal in that it accounts consistently for
the expectations and minimizations in these two stages, but no anticipation beyond that.

The principle of
two-step look-ahead is to select the next observation location $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
we were to (optimally) add one more observation after that again. In order to formalize this concept, we must
extend the notation for EIBV in the future (after observation
$\stage + 1$ has been made). We let $\currentEIBV(\cdot; u, y)$
denote the EIBV where expectations are taken conditional on the data
available at stage $n$ and on an additional observation $y$ at $u$ at stage $n+1$.

\begin{criterion}[2-step look-ahead]
      The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
      \begin{align}\label{critLA}
          C_{\text{2-steps}}(u) &= \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
                  \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,
      Y\right)\right]
      \end{align}
      where $Y$ is the random data realization of $\gp_{\spatloc}$ according to its
      conditional law at step $\stage$ with the
      dependence of the set of candidates on the current location
      having been made explicit for the second stage of measurements.
    \end{criterion}

    In a practical setting, the first expectation can be computed by
    Monte Carlo sampling of data $Y$ from its conditional
    distribution. For each of these data samples, the second
    expectation is solved using the closed-form expressions for EIBV
    provided by Proposition \ref{propo2}, now with conditioning on the
    first stage data already going into the co-Kriging updating
    equations.

\subsection{Simulation study}
\label{sec:simulations}

\subsubsection{Static and Sequential Sampling Designs}
\label{sec:sampling_designs}

%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
%\end{figure}
%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.

We compare three different static designs denoted
\textit{static\_north}, \textit{static\_east}, and
\textit{static\_zigzag} with the three described sequential approaches
\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
AUV sampling paths are pre-scripted and cannot be altered.
For a fixed survey length, a closed-form expression for the EIBV is
available as in Proposition \ref{propo_eibv}. However, for the sequential
approaches this is not the case. For comparison, the properties are
therefore evaluated using Monte Carlo sampling over several replicates
of realizations from the model while conducting simulated sequential
surveys for each one. We also compare predictive
performance measured by root mean square error (RMSE) for temperature
and salinity estimates as well as the variance reduction in these
two variables. It is important to note that the objective function
used by the AUV is focused on reducing the
EIBV, but we nevertheless expect that we will achieve good predictive
performance for criteria such as RMSE as well. Another non-statistical
criterion that is important for practical purposes is the computational
time needed for the strategy.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
%\caption{The equilateral waypoint graph used to discretize the
%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
%\label{fig:wp_graph}
%\end{figure}

\begin{figure}[!b] 
\centering 
\subfigure[The waypoint graph.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
\hfill
\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
\label{fig:wp_graph}
\end{figure}

Each strategy is conducted on an equilateral grid as shown in
Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at the southern end of the domain (node 53). It then moves
along edges in the waypoint graph while collecting data which are
assimilated onboard to update the GRF model. This is used in the evaluation
of the next node to sample.  The procedure is run for $10$ stages. A
total of 100 replicate simulations were conducted with all strategies.

\subsection{Simulation Results}

The results of the replicate runs are shown in
Fig.~\ref{fig:sim_results}, where the different criteria are plotted
as a function of survey distance. Fig.~\ref{fig:avg_ev} shows the
resulting drop in realized IBV for each of the six strategies. IBV
reduction is largest for the \textit{myopic} and
\textit{look-ahead} strategies, each performing almost equally; this
is expected as the two criteria (Eq. \eqref{critSEQ} and
\eqref{critLA}) are sensitive to differences in IBV. The
\textit{static\_north} design also does well here because the path is
parallel to the boundary between the water masses.

\begin{figure}[!h]
  \centering
  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \hfill \subfigure[RMSE between estimated field and
  truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
  \hfill \subfigure[Explained variance
  $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
  \hfill \subfigure[Computational time for inference (the lines for \textit{naive}, \textit{static\_north}, \textit{static\_east}, and \textit{static\_zigzag}
  overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
\caption{Simulation results from 100 replicate simulations for 10 sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
\label{fig:sim_results}
\end{figure}

Fig.~\ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
in RMSE and increase in explained variance, respectively. Both
\textit{myopic} and \textit{look-ahead} strategies perform well here,
but some of the \textit{static\_east} and \textit{static\_zigzag} also
achieve good results because they cover large parts of the domain
without re-visitation. Sequential strategies targeting IBV will
sometimes not reach similar coverage, as interesting data may draw the
AUV into twists and turns. There is a relatively large variety in the
replicate results as indicated by the vertical lines. Nevertheless,
the ordering of strategies is similar.

Fig.~\ref{fig:avg_time} shows the computational effort: the
\textit{naive} strategy is on par with the static designs, while the
\textit{myopic} strategy is slower because it evaluates expected
values for all candidate directions at the waypoints. But it is still
able to do so in reasonable time, which allows for real-world
applicability. The \textit{look-ahead} strategy is much slower, reaching levels that
are nearly impractical for execution on an AUV. Some pruning of the
graph is performed to improve the performance, such as ruling out
repeated visitations. Further pruning of branches or inclusion of
other heuristics could be included for better performance. Then again,
the inclusion of such heuristics is likely a contributing factor for
the \textit{look-ahead} strategy failing to outperform the
\textit{myopic} strategy.

%In Fig. \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown. The
%\textit{naive} strategy often gets stuck in the southern part of the
%domain because it is too focused on the probabilities near $0.5$. The
%\textit{myopic} strategy covers a wider domain than the naive or
%look-ahead. There are several reasons for this. First, a greedy
%approach will tend to put more emphasis on promising locations close
%to the agent, which may lead away from the centre. Second, as the
%agent evaluates the impact of locations further away (look-ahead)
%where assimilated data has less predictive power, the GP model (which
%is centered here) will act to restrict paths deviating from the
%central zone.

We studied the sensitivity of the results by modifying the input
parameters to have different correlations between temperature and
salinity, standard deviations, and spatial correlation range.  In all
runs, the \textit{myopic} and \textit{look-ahead} strategies perform
the best in terms of realized IBV, and much better than
\textit{naive}. The \textit{look-ahead} strategy seems to be
substantially better than the \textit{myopic} design only for very
small initial standard deviations or very large spatial correlation
range. 
%\textit{static\_north} continues to be the best static design
%for IBV, while \textit{static\_zigzag} is the best design for the
%other predictive performance measures, especially so with large
%spatial correlation range. 
We also ran simulation studies with only temperature data, and for
realistic correlation levels between temperature and salinity, the IBV
results are not much worse when only temperature data are
available. In addition to the comparison made in
Table~\ref{tab:sim_rhoab}, the current setting includes spatial
correlation and this likely strengthen the effect of temperature information. However, it seems that having temperature data alone
does a substantially worse job in terms of explained variance.

\section{Case Study - Mapping a River Plume}
\label{sec:case_study}

To demonstrate the applicability of using multivariate EPs and the IBV
to inform oceanographic sampling, we present a case study mapping a
river plume with an AUV. The experiment was performed in Trondheim,
Norway, surveying the Nidelva river (Fig.~\ref{fig:nidelven}). The
experiments were conducted in late Spring 2019, when there is still
snow melting in the surrounding mountains so that the river water is colder than the water in the fjord. The experiment was
focused along the frontal zone that runs more or less parallel to the
eastern shore.

\subsection{Model Specification}
\label{sec:exp_modeling}

The statistical model parameters were specified based on a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions and correlation
structures. Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the trend by linear regression, where both temperature and salinity were assumed to increase linearly with the west coordinate. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
%\begin{figure}[!h] 
% \centering 
%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
%\caption{Data analysis from a preliminary trial experiment using the
%  AUV. Left: Residual plot of temperature and salinity along with
%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
%  distribution with two degrees of freedom. Right: Empirical variogram
%  of the salinity and temperature data.} \label{fig:parest}
%\end{figure}
\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
  \hfill
  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
  \caption{Data analysis from a preliminary trial experiment using the
    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
    salinity and temperature data.}
\label{fig:parest}
\end{figure}
Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates larger variability in salinity than temperature,
and a positive correlation ($0.5$) between the two variables. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The modeled and
theoretical curves are rather similar, which indicates that the Gaussian
model with constant spatial variance and correlation fits reasonably well. Fig.~\ref{fig:parest_c} shows the
empirical variogram of the residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
after about $150$ m.  The working assumption of a separable covariance
function is hence not unreasonable.

Based on the analysis in Fig.~\ref{fig:parest}, the resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean. The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water. Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.

\begin{table}[!h]
\centering
\begin{tabular}{lrr}
\toprule
Parameter & Value & Source\\
\midrule
\rowcolor{Gray}
Cross correlation temperature and salinity & 0.5 & AUV observations\\
Temperature variance &  0.20 & AUV observations (variogram)\\
\rowcolor{Gray}
Salinity variance &  5.76 & AUV observations (variogram)\\
Correlation range  & 0.15 km & AUV observations (variogram)\\
\rowcolor{Gray}
River temperature  & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
Ocean temperature $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
\rowcolor{Gray}
River salinity $S_{river}$ & $14.0$ g/kg & AUV observations\\
Ocean salinity $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
\rowcolor{Gray}
Threshold in temperature & $10.5\,^{\circ}\mathrm{C}$ & User specified \\
Threshold in salinity & $18.0$ g/kg & User specified \\
\rowcolor{Gray}
\bottomrule
\end{tabular}
\caption{Model and threshold parameters from an initial AUV
  survey. Observations were taken across the front while crossing from
  fresh, cold river water to saline and warmer ocean waters.}
\label{tab:experiment_param}
\end{table}

\subsection{Experimental Setup}

A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
sensor was used to provide salinity and temperature measurements.  The
AUV is a powered untethered platform that operates at $1$-$3$ m/s in
the upper water column. It has a multicore GPU NVIDIA Jetson TX1
(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
was built on top of the autonomous Teleo-Reactive EXecutive
(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
that the measurements are conditionally independent because the
salinity is extracted from the conductivity sensor which is different
from the temperature sensor. We specify variance $0.25^2$ for both
errors, which is based on a middle ground between the nugget effect in
the empirical variogram and the sensor specifications.

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater
  Vehicle (LAUV) platform for upper water-column exploration used in
  our experiments.}
\label{fig:lauv}
\end{figure}

The AUV was running a \textit{myopic} strategy to decide between
sampling locations on the waypoint graph distributed over an
equilateral grid, as shown in the grey-colored lattice in
Fig.~\ref{fig:map}.  At each stage, it takes the AUV about 30 seconds
to assimilate data and evaluate the EIBV for all the possible waypoint alternatives.  It
was set to start in the south-center part of the waypoint graph. A
survey was set to take approximately 40 minutes, visiting 15 waypoints
on the grid, with the vehicle running near the surface to capture the
plume. On its path from one waypoint to the next, the AUV collects
data with an update frequency of 30 seconds, giving three measurements
per batch in the updates at each stage.

\begin{figure*}[!h]
\centering
\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
\hspace{0.3cm}
\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
\caption{Results from mapping the Nidelva river, Trondheim, Norway
  over two survey missions. \ref{fig:map} shows an overview of the
  survey area overlaid with the AUV path in black and dashed
  line. Note the shaded region indicating a typical frontal
  region. \ref{fig:res_both} shows the collected temperature data as
  colored trails. Note waypoint 5 (WP5) which indicates where the two
  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
  collected salinity data overlaid on the final EP, which indicate the
  AUVs statistical impression of the front. For both missions the
  temperature and salinity data correspond with an indication of the
  EP front. About 2 hours time separated the two runs.}
\label{fig:results}
\end{figure*}

\subsection{Results}

Two survey missions (1 and 2), were run successively, with a short
break in between. The resulting path of the selected waypoints are
shown in the map in Fig.~\ref{fig:map}, both within the expected
frontal region (shaded pink). The recorded temperatures are shown as
colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
temperature difference between fjord and riverine waters. The salinity
data are then shown separately, overlaid with the estimated EP for
each survey in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2}.

Both surveys successfully estimated and navigated the separation zone,
crossing the frontal boundary multiple times. As conditions changed
slightly between the two surveys, the resulting trajectory (after
waypoint 5) is shown to deviate. Survey 1 continued northwards,
tracking the north-eastern portion of the front, while Survey 2 turned
west, mapping the south-western region.

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner.

\section{Closing remarks}
\label{sec:concl_disc}

This work builds on a multidisciplinary effort combining statistical
methods with robotic surveying techniques for oceanographic
applications. We show how observation practices can gain efficiency
and accuracy from the development of statistical techniques for
spatial monitoring. We further demonstrate the opportunities available
for real-time multivariate spatial data gathering and analysis on
autonomous platforms, which statisticians can exploit to create
general-purpose toolkits for similar applications.

In particular, we derive and show results for characterizing phenomena
connected to the properties of water masses. The characterization of
uncertainties in random sets is extended in the vector-valued case with new results for the
expected integrated Bernoulli variance reduction achieved by spatial
sampling designs. This is provided in semi-analytical form for the situation
with a static design, and then extended to the adaptive situation. The
sequential derivations provide new insights into efficient
applications of adaptive data collection, as demonstrated in our
application.

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but the operation must
likely approximate calculations by concentrating numerical integration efforts on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. We did not
consider any temporal effects, which would be relevant on a larger
time scale. We consider the extension to spatio-temporal modeling as
future work, and envision that advection-diffusion equations could be
useful in this kind of modeling
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could still be run onboard and augmented by dynamical
models. If high-resolution ocean model or satellite data are available, it is hopefully possible to fit such more complex statistical models.

The spatial-statistical design criterion building on random sets is
relevant in our setting with different water properties.  We show
mathematical generality beyond the expected integrated Bernoulli
variance, for instance that of volume uncertainties which is possibly
more relevant, but requires more computational resources.  Such
criteria could be particularly useful in other oceanographic settings
related to mapping of algal-blooms, anoxic zones or open water fronts
\cite{costa19}.  Of course, other criteria are also relevant for
instance, hybrid or multi-attribute criteria that could balance goals
of exploration and exploitation in this situation. Equally, such
techniques have significant use cases in downstream decision-making,
with policy makers and regulators who need to make difficult decisions
related to fish farming or other marine resource operations. Value of
information analysis \citep{Eidsvik:15} could be used to evaluate
whether information is likely to result in improved decision-making,
in such a context.  We also foresee opportunities related to design of
experiments for multivariate processes using our notion of generalized
locations. For instance, this is likely to be useful for
internet-of-things applications or computer experiments where some
observations or evaluations are rather inexpensive, while others 
must only be done when they are really valuable.

In our context the myopic strategy performs well, and for
computational reasons we did not go in depth on the dynamic program
solution. There has been much creative work on finite horizon
optimization in the robotics literature including probabilistic road
maps and rapidly-exploring random trees \citep{karaman2011sampling},
but their statistical properties are not yet clear.  
In some cases it is also limiting to use the waypoint graph, and it would be beneficial to allow more continuous updates and navigation at the highest frequency allowed by the AUV.
It is equally
interesting to explore the additional flexibility that can be gained
by having multiple vehicles co-temporally exploring a spatial or
spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
would enable concurrent sampling in different parts of the space, or
opportunities to move in parallel to best capture the excursion set.
The sometimes conflicting topics of autonomy and communication have
caught much interest in control engineering lately
\citep{zolich2019survey}, and statistical evaluations could contribute
in the future. In our context, autonomy is highlighted, and this is
reflected in our modeling and computational assumptions. With the
possibilities of wifi communication to a mother-ship at waypoints,
more advanced modeling and computational routines are of course
possible, but with new challenges related to design and planning for
when and where the communication should occur.

\section*{Acknowledgements}

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
support from the Swiss National Science Foundation, project number
178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to aknowledge support of Idiap Research Institute, where he had is main affiliation when the first version of this work was submitted. The authors would also like to thank Niklas Linde (University of Lausanne) for pointing out the relevance of revisiting the proposed approach for estimating volumes of mixing zones, e.g. in hydrology.

%\begin{supplement}
%\sname{Supplement A}\label{suppA}
%\stitle{Title of the Supplement A}
%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
%\sdescription{Dum esset rex in
%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
%\end{supplement}

% == Adding references
\footnotesize
\bibliographystyle{imsart-nameyear}
\bibliography{ref}

\section*{Appendix}
\input{appendix}

% AOS,AOAS: If there are supplements please fill:
%\begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{Some text}
%\end{supplement}

% === Not used
% After reviews: - This is moved to Section 2.1.

%Traditional data collection at sea has typically been based on static
%buoys, Lagrangian floats, or ship-based methods, with significant
%logistical limitations that directly impact coverage and sampling
%resolution. Modern methods using satellite remote-sensing provide
%large-scale coverage but have limited resolution, are limited to
%sensing the surface of the ocean, and are impacted by cloud cover. The
%advent of robust mobile robotic platforms \citep{Bellingham07} has
%resulted in significant contributions to environmental monitoring and
%sampling. In particular, autonomous underwater vehicles (AUVs), have
%advanced the state of sampling and consequently have made robotics an
%integral part of ocean observation; our previous work has contributed
%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
%focusing on hierarchical statistical models; \cite{sahu2008space},
%studying spatio-temporal models for sea surface temperature and
%salinity data; and \cite{mellucci2018oceanic} looking at the
%statistical prediction of features using an underwater glider.

%\begin{itemize}
%
%\item Full numerical ocean models cannot provide accurate results
%  online if run on robotic sensing platforms, as onboard computers
%  cannot deliver the computational power required. Hence statistical
%  proxy models of the environment must be used for learning where to sample.
%
%\item With limited available information about the state of the ocean, there is substantial value in reacting to
%  information obtained from measurements taken in-situ. This
%  acquired information must be assimilated into statistical models that
%  can be used to inform decisions on where to sample
%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
%
%\item For sampling problems related to environmental sensing, the
%  number of choices (i.e. locations, trajectories, and candidate
%  designs) is enormous, creating a
%  trade-off between optimization (finding the most resource-efficient
%  design to collect necessary data) and computability (arriving at a
%  solution in reasonable time). To successfully resolve features, this
%  trade-off has to be considered in development and practice.
%
%\end{itemize}

%Addressing this, the combination of statistical tools and robotic platforms is a
%natural symbiosis which enables information-based sensing. Central to
%this is the ability to model spatially-correlated variables and
%provide formal measures of uncertainty. Our formulation is based on
%Gaussian Processes (GPs) as they allow efficient implementation 
%and evaluation in real time onboard a robotic platform.

%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.

%While the focus has often been on
%biological and anthropogenic impact from micro-plastics to
%pollution, biological oceanographers have focused intently on studying
%micro-organisms at the base of the human food web. These organisms are
%critically impacted by the changing dynamics in the upper water-column,
%especially in coastal zones which are complex and often hard to observe
%in space and time. By studying the bio-geochemical processes in the
%upper water-column scientists can measure the impact of change, natural
%or anthropomorphic, and provide an informed opinion to policy makers to
%effect changes in preserving the environment. However, the challenge of 
% The pressure on marine resources is growing and increased accuracy,
% resolution, and persistent monitoring of the oceans is crucial for
% long-term sustainable management.

% A
% sustained focus on prioritized and efficient data collection strategies
% have therefore started to emerge. The advent of marine robotic
% platforms, especially

% provided means to execute this prioritization through the capacity of
% autonomy and data-driven sampling, where data collection in principle
% can be optimized. These capabilities have made

% of the emerging sensing practice for ocean science, allowing scientists
% to increase the observational efficiency and resolution beyond what was
% previously possible. But how should a robotic platform, such as an AUV,
% effectively prioritize and identify important regions for sampling? The
% answer to this question relates to

% , and
% the application domain is clearly an arena where statisticians can
% contribute.
% There has recently been some statistical attention in oceanography:

%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest. Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity. One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}. These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively. Quantification of gradient features is therefore a much needed competence in robotic sampling.
\end{document}